---
title: "02_vireo_global"
author: "jibsch"
date: "2022-06-30"
output: workflowr::wflow_html
editor_options:
  chunk_output_type: inline
---



```{r}
library(Seurat)
library(tidyverse)
library(RColorBrewer)
library(ggrepel)
source("~/tools/BioinfTools/Plotting_Functions.R")
```

```{r}
merged <- readRDS("~/projects/Lewinlab_AMC/AMC/Analysis/output/data/2022-07-20_AMC_merged.rds")

vireo = read_tsv("../mergeBams/vireo//donor_ids.tsv")
vireo.01 = read_tsv("../mergeBams/vireo_0.01//donor_ids.tsv")
vireo.05 = read_tsv("../mergeBams/vireo_0.05//donor_ids.tsv")
vireo.005 = read_tsv("../mergeBams/vireo_0.005//donor_ids.tsv")
vireo.0025 = read_tsv("../mergeBams/vireo_0.0025//donor_ids.tsv")
vireo.001 = read_tsv("../mergeBams/vireo_0.001//donor_ids.tsv")

vireo.01$cell = gsub("-1","",vireo.01$cell)
row.names(vireo.01) = vireo.01$cell

merged$donor_1pct = vireo.01[row.names(merged@meta.data),]$donor_id


table(merged$donor_1pct, merged$orig.ident)

table(merged$donor_new, merged$hashedDrops_classification)

table(merged$donor_1pct, merged$HTO_classification)

```
## Sample 4A

```{r}
vir_local = read_tsv("~/projects/Lewinlab_AMC/AMC/AMC_extractedData/vireosnp_output_merged/Sample_4A/donor_ids.tsv")
head(vir_local)
vir_local$cell = gsub("^","4A_",vir_local$cell)

vir_local_3A = read_tsv("~/projects/Lewinlab_AMC/AMC/AMC_extractedData/vireosnp_output_merged/Sample_3A/donor_ids.tsv")
head(vir_local)
vir_local_3A$cell = gsub("^","3A_",vir_local_3A$cell)

vir_local_3B = read_tsv("~/projects/Lewinlab_AMC/AMC/AMC_extractedData/vireosnp_output_merged/Sample_3B/donor_ids.tsv")
head(vir_local)
vir_local_3B$cell = gsub("^","3B_",vir_local_3B$cell)

summary(vir_local$cell %in% vireo$cell)

vireo.01 %>% inner_join(vir_local, by = "cell") -> x

table(x$donor_id.x, x$donor_id.y)
```

```{r}
lapply(list(vireo, vireo.01, vireo.05), function(x)table(x$donor_id))
```
```{r}
calls = list(vireo,vireo.001, vireo.0025, vireo.005,vireo.01, vireo.05)
names = c("MAF=0","MAF=0.1%", "MAF=0.25%","MAF=0.5%", "MAF=1%", "MAF=5%")
names(calls) = names
ps = list()
for(i in 1:(length(calls)-1)){
  for(j in (i+1):length(calls)){
    data.frame(vireo_0 = calls[[i]]$donor_id, vireo_01 = calls[[j]]$donor_id) %>%
    filter(!vireo_0 %in% c("doublet", "unassigned") & !vireo_01 %in% c("doublet", "unassigned")) %>%
    group_by(vireo_0, vireo_01) %>% summarise(n = n()) %>%
    #group_by(vireo_01) %>% 
    ungroup() %>%
   # group_by(vireo_0) %>%
    mutate(r1 = min_rank(n)) %>% 
    # ungroup() %>% group_by(vireo_01) %>%
    # mutate(r2 = min_rank(n)) %>% head()
    # 
    #mutate(s = ifelse(n() > 1,scale(n)[r], 0.5)) %>% 
    ggplot(aes(reorder(vireo_0,r1), reorder(vireo_01,r1))) +
    geom_tile(aes(fill = log10(n))) +
    geom_text(aes(label = n)) +
    labs(x=names[i], y=names[j])+
    theme_classic() +
        theme(axis.text.x = element_text(angle=90, vjust  =0.5), legend.position = "none")-> p
    ps[[paste(names[i],"vs",names[j])]] = p
  }
}


ps2 = mapply(function(x,z){
  x %>% inner_join(vir_local, by = "cell") %>%
    group_by(donor_id.x, donor_id.y) %>% summarise(n = n())%>%
    ungroup() %>% mutate(r=min_rank(n)) %>%
    ggplot(aes(reorder(donor_id.x,r), reorder(donor_id.y,r))) +
    geom_tile(aes(fill = log10(n))) +
    geom_text(aes(label = n, color = n>15)) +
    labs(x=z, y="Sample4A")+
    theme_classic() +
        theme(axis.text.x = element_text(angle=90, vjust  =0.5), legend.position = "none") +
    scale_color_manual(values = c("white", "black"))
}, calls, names, SIMPLIFY = F)

ps3 = mapply(function(x,z){
  x %>% inner_join(vir_local_3A, by = "cell") %>%
    group_by(donor_id.x, donor_id.y) %>% summarise(n = n())%>%
    ungroup() %>% mutate(r=min_rank(n)) %>%
    ggplot(aes(reorder(donor_id.x,r), reorder(donor_id.y,r))) +
    geom_tile(aes(fill = log10(n))) +
    geom_text(aes(label = n, color = n>15)) +
    labs(x=z, y="Sample3A")+
    theme_classic() +
        theme(axis.text.x = element_text(angle=90, vjust  =0.5), legend.position = "none") +
    scale_color_manual(values = c("white", "black"))
}, calls, names, SIMPLIFY = F)

ps4 = mapply(function(x,z){
  x %>% inner_join(vir_local_3B, by = "cell") %>%
    group_by(donor_id.x, donor_id.y) %>% summarise(n = n())%>%
    ungroup() %>% mutate(r=min_rank(n)) %>%
    ggplot(aes(reorder(donor_id.x,r), reorder(donor_id.y,r))) +
    geom_tile(aes(fill = log10(n))) +
    geom_text(aes(label = n, color = n>15)) +
    labs(x=z, y="Sample3B")+
    theme_classic() +
        theme(axis.text.x = element_text(angle=90, vjust  =0.5), legend.position = "none") +
    scale_color_manual(values = c("white", "black"))
}, calls, names, SIMPLIFY = F)

```
```{r}
table(vireo.01$best_doublet[vireo.01$best_singlet == "donor3"])
```

##Compare Global and Local Assignments

```{r}
summary(merged$HTO_classification[merged$donor_1pct %in% paste0("donor",c(0,11,3,4,6))] != "Negative")
table(merged$HTO_classification[merged$donor_1pct %in% paste0("donor",c(0,11,3,4,6))]) %>% as.data.frame() %>%
  filter(Var1!= "Negative") %>% 
  ggplot(aes(reorder(Var1, Freq), Freq)) + geom_bar(stat="identity") +
  theme_classic() + labs(x= "#AB")
```


```{r}
merged@meta.data %>% group_by( donor_1pct,HTO_classification) %>%
  summarise(n=n()) %>%
  mutate(r=min_rank(n)) %>%
  ggplot(aes(donor_1pct, HTO_classification)) +
  geom_tile(aes(fill = log10(n)))+
  geom_text(aes(label=n, colour = n>30)) +
  theme_classic() +
   scale_color_manual(values = c("white", "black")) +
  theme(legend.position = "none", axis.text.x = element_text(angle=90, vjust  =0.5))
```


```{r}
merged@meta.data %>% group_by( donor_1pct,HTO_classification, orig.ident) %>%
  summarise(n=n()) %>%
  mutate(r=min_rank(n)) %>%
  ggplot(aes(donor_1pct, HTO_classification)) +
  geom_tile(aes(fill = log10(n)))+
  geom_text(aes(label=n, colour = n>30)) +
  theme_classic() +
  facet_wrap(~orig.ident)+
   scale_color_manual(values = c("white", "black")) +
  theme(legend.position = "none", axis.text.x = element_text(angle=90, vjust  =0.5))
```

```{r}
ps_1pct = lapply(list("1B", "2A", "2B", "3A", "3B", "4A"), function(x){
  vir_local = read_tsv(paste0("~/projects/Lewinlab_AMC/AMC/AMC_extractedData/vireosnp_output_merged/Sample_",x,"/donor_ids.tsv"))
  vir_local$cell = gsub("^",paste0(x,"_"),vir_local$cell)
  vir_local$cell = gsub("-1","",vir_local$cell)
  head(vir_local)
  vireo.01 %>% inner_join(vir_local, by = "cell") %>%
    group_by(donor_id.x, donor_id.y) %>% summarise(n = n())%>%
    ungroup() %>% mutate(r=min_rank(n)) %>%
    ggplot(aes(reorder(donor_id.x,r), reorder(donor_id.y,r))) +
    geom_tile(aes(fill = log10(n))) +
    geom_text(aes(label = n, color = n>15)) +
    labs(x="Global vireo (1% MAF)", y=paste0("Local vireo ",x," (10% MAF)"))+
    theme_classic() +
        theme(axis.text.x = element_text(angle=90, vjust  =0.5), legend.position = "none") +
    scale_color_manual(values = c("white", "black"))
})
```
```{r}
vir_local_2B = read_tsv("~/projects/Lewinlab_AMC/AMC/AMC_extractedData/vireosnp_output_merged/Sample_2B/donor_ids.tsv")
vir_local_2B$cell = gsub("-1","",vir_local_2B$cell) %>% gsub("^", "2B_",.)
row.names(vir_local_2B) = vir_local_2B$cell

merged$donor_2B = vir_local_2B[row.names(merged@meta.data),]$donor_id


merged@meta.data %>% filter(orig.ident == "Sample_2B") %>% group_by( donor_2B,HTO_classification) %>%
  summarise(n=n()) %>%
  mutate(r=min_rank(n)) %>%
  ggplot(aes(donor_2B, HTO_classification)) +
  geom_tile(aes(fill = log10(n)))+
  geom_text(aes(label=n, colour = n>30)) +
  theme_classic() +
   scale_color_manual(values = c("white", "black")) +
  theme(legend.position = "none", axis.text.x = element_text(angle=90, vjust  =0.5))
```

```{r}
vir_local_3B = read_tsv("~/projects/Lewinlab_AMC/AMC/AMC_extractedData/vireosnp_output_merged/Sample_3B/donor_ids.tsv")
vir_local_3B$cell = gsub("-1","",vir_local_3B$cell) %>% gsub("^", "3B_",.)
row.names(vir_local_3B) = vir_local_3B$cell

merged$donor_3B = vir_local_3B[row.names(merged@meta.data),]$donor_id


merged@meta.data %>% filter(orig.ident == "Sample_3B") %>% group_by( donor_3B,HTO_classification) %>%
  summarise(n=n()) %>%
  mutate(r=min_rank(n)) %>%
  ggplot(aes(donor_3B, HTO_classification)) +
  geom_tile(aes(fill = log10(n)))+
  geom_text(aes(label=n, colour = n>30)) +
  theme_classic() +
   scale_color_manual(values = c("white", "black")) +
  theme(legend.position = "none", axis.text.x = element_text(angle=90, vjust  =0.5))
```

### New Comparison

```{r}
as.data.frame(table(merged$sample)) %>% mutate(Var1 = gsub(" ","_",Var1)) %>% inner_join(as.data.frame(table(merged2$Donor_ID_conservative)), by = "Var1") %>% pivot_longer(-Var1) %>% ggplot(aes(Var1, value)) + 
  geom_bar(stat="identity", aes(fill = name), position= "dodge", colour = "black") + theme_classic() +
  scale_fill_manual(labels = c("Hybrid global", "Conservative local"), values = RColorBrewer::brewer.pal(2, "Set2")) +
  theme(legend.position = "top", axis.text.x = element_text(angle=45, hjust = 1)) +
  labs(x = element_blank(), y = "# cells", fill = "Strategy:")

as.data.frame(table(merged$sample)) %>% mutate(Var1 = gsub("Donor ","donor",Var1)) %>% inner_join(as.data.frame(table(merged2$latent_donor)), by = "Var1") %>% 
  pivot_longer(-Var1) %>% filter(Var1 != "doublet") %>%
  ggplot(aes(Var1, value)) + 
  geom_bar(stat="identity", aes(fill = name), position= "dodge", colour = "black") + theme_classic() +
  scale_fill_manual(labels = c("Hybrid global", "Conservative local"), values = RColorBrewer::brewer.pal(2, "Set1")) +
  theme(legend.position = "top", axis.text.x = element_text(angle=45, hjust = 1)) +
  labs(x = element_blank(), y = "# cells", fill = "Strategy:")

as.data.frame(table(merged$sample)) %>% mutate(Var1 = gsub(" ","_",Var1)) %>% 
  inner_join(as.data.frame(table(merged2$Donor_ID_conservative)), by = "Var1") %>% 
  inner_join(as.data.frame(table(merged2$Donor_ID_HTOdemux_localSNP)), by = "Var1") %>% 
  mutate(Var1 = gsub("Donor_","donor",Var1)) %>%
  inner_join(as.data.frame(table(merged2$latent_donor)), by = "Var1") %>%
  pivot_longer(-Var1) %>% ggplot(aes(Var1, value)) + 
  geom_bar(stat="identity", aes(fill = name), position= "dodge", colour = "black") + theme_classic() +
  scale_fill_manual(labels = c( "Hybrid global", "Local", "Conservative local", "Latent class"), values = RColorBrewer::brewer.pal(4, "Set2")) +
  theme(legend.position = "top", axis.text.x = element_text(angle=45, hjust = 1)) +
  labs(x = element_blank(), y = "# cells", fill = "Strategy:")

```




## Hybrid Assignment
```{r}
merged$donor_hybrid = merged$donor_1pct
merged$donor_hybrid = ifelse(merged$donor_id_snp == "donor5" & merged$orig.ident == "Sample_4A", "donor90", merged$donor_hybrid)
merged$donor_hybrid = ifelse(merged$donor_id_snp == "donor3" & merged$orig.ident == "Sample_4A", "donor3", merged$donor_hybrid)

merged$sample = recode(merged$donor_hybrid,
                       "donor0" = "PBMC C",
                       "donor1" = "Donor 49",
                       "donor3" = "PBMC D",
                       "donor4" = "PBMC E",
                       "donor5" = "Donor 38",
                       "donor6" = "PBMC B",
                       "donor7" = "Donor 54",
                       "donor8" = "Donor 57",
                       "donor9" = "Donor 48",
                       "donor10" = "Donor 62",
                       "donor11" = "PBMC F",
                       "donor12" = "Donor 41",
                       "donor13" = "Donor 64",
                       "donor14" = "Donor 56",
                       "donor90" = "PBMC G"
                       )

table(merged$sample)

```


```{r}
assignment_table = read.table(text = "Sample 	Donor 38	Donor 41	Donor 48	Donor 49	Donor 54	Donor 56	Donor 62	Donor 64	Donor 57
Sample_1B 	C01D+1 	NA 	C13D+1 	C01D+8 	C22D+1 	C01D+1 	Baseline 	C01D+8 	NA 
Sample_2A 	C01D+8 	C04D-1 	C13D-1 	C04D-1 	C04D-1 	C01D+8 	C04D-1 	Baseline 	Baseline 
Sample_2B 	C22D+1 	C13D+1 	C04D+1 	Baseline 	C04D+1 	C04D+1 	C01D+8 	C04D-1 	C01D+1 
Sample_3A 	Baseline 	Baseline 	C22D+1 	C13D-1 	C13D+1 	C04D-1 	NA 	C01D+1 	C01D+8 
Sample_3B 	NA 	C22D+1 	C04D+1 	C13D+1 	C13D-1 	Baseline 	C04D+1 	C04D+1 	C04D-1 
Sample_4A 	C01D+8 (Donor41) 	C13D-1 	C04D+1 	C04D+1 	C04D+1 	Baseline 	C04D+1 	C04D+1 	C04D+1", sep = "\t", header= T)

assignment_table %>% pivot_longer(-Sample, names_to = "sample", values_to = "condition") %>%
  mutate(sample = gsub("[.]", " ", sample),
         Sample = gsub(" $","",Sample),
         condition = gsub(" $","", condition)) %>%
  rename("orig.ident" = "Sample")-> assignments

merged@meta.data %>% left_join(assignments) -> x
x$condition = ifelse(grepl("PBMC", x$sample), "PBMC", x$condition)
x$condition[x$condition == "NA"] = NA

x$condition[x$orig.ident == "Sample_4A" & x$sample == "Donor 41" & x$HTO_classification == "HTO-1"] = "C01D+8"
x$condition[x$orig.ident == "Sample_4A" & x$sample == "Donor 41" & !x$HTO_classification %in% c("HTO-1","HTO-2")] = NA
x$condition[x$orig.ident == "Sample_1B" & x$sample == "Donor 41"] = "PBMC"

table(x$orig.ident,x$condition)

merged$condition = x$condition
merged$sample[merged$orig.ident == "Sample_1B" & merged$sample == "Donor 41"] = "PBMC A"
merged$condition[merged$orig.ident == "Sample_2A" & merged$sample == "PBMC D"] = NA

DimPlot(merged, group.by = "condition", label = T)


```
### No PBMC
```{r}
merged2 = merged[,!is.na(merged$condition) ]
merged2 = SCTransform(merged2)
varfeatures = merged2@assays$SCT@var.features[!grepl("TR[AB]", merged2@assays$SCT@var.features)]
merged2 = RunPCA(merged2, features = varfeatures)
merged2 = RunUMAP(merged2, dims = 1:25)

DimPlot(merged2, group.by = c("sample", "condition"), label = T)
```
```{r}
merged3 = merged2[,merged2$condition != "PBMC"]
merged3 = SCTransform(merged3)

varfeatures = merged3@assays$SCT@var.features[!grepl("TR[AB]", merged3@assays$SCT@var.features)]
merged3 = RunPCA(merged3, features = varfeatures)
merged3 = RunUMAP(merged3, dims = 1:25)

DimPlot(merged3, group.by = c("sample", "orig.ident", "condition"), label = T)

merged3 = FindNeighbors(merged3, dims = 1:25)
merged3 = FindClusters(merged3)
DimPlot(merged3, label = T)
```

```{r}
markers = FindAllMarkers(merged3, only.pos = F)

write.csv(markers, "../output/markers_cd8s.csv", quote = F, row.names = F)

markers %>% filter(cluster %in% c(7:9)) %>%
  group_by(cluster) %>% slice_head(n=5)
```
```{r}
library(org.Hs.eg.db)
markers$entrez = mapIds(org.Hs.eg.db, markers$gene, "SYMBOL", column = "ENTREZID")
keggs = lapply(split(markers$entrez, markers$cluster), function(x) kegga(species = "Hs", de = x))
```

### Signatures

```{r}
naive = list(plus = c("SELL", "TCF7", "CCR7", "IL17RA", "S1PR1", "KLF2"),
             minus = c("KLRG1","CX3CR1","S1PR5","TNF","IFNG","GZMB","PRF1","MKI67","PDCD1", "HAVCR2", "CTLA4"))
merged3 = AddModuleScore(merged3, features = naive)
names(merged3@meta.data)[names(merged3@meta.data) == c("Cluster1", "Cluster2")] = c("Naive_pos", "Naive_neg")
# merged3$Naive_neg = -merged3$Naive_neg
merged3$Naive = scale(merged3$Naive_pos) - scale(merged3$Naive_neg)
ggarrange(plots = list(plotto_signature_scoring_plot("Naive_pos", merged3, size = 1),
                       plotto_signature_scoring_plot("Naive_neg", merged3, size =1),
                       plotto_signature_scoring_plot("Naive", merged3, size =1)))
```
```{r}
effector = list(plus =c("KLRG1", "CX3CR1", "S1PR5", "TNF", "IFNG", "GZMB", "PRF1"),
                neg = c("SELL", "TCF7", "LEF1", "CCR7", "IL17RA"))

merged3 = AddModuleScore(merged3, features = effector)
names(merged3@meta.data)[c(length(names(merged3@meta.data)) -1, length(names(merged3@meta.data)))] = c("Effector_pos", "Effector_neg")
merged3$Effector = scale(merged3$Effector_pos) - scale(merged3$Effector_neg)
ggarrange(plots = list(plotto_signature_scoring_plot("Effector_pos", merged3, size = 1),
                       plotto_signature_scoring_plot("Effector_neg", merged3, size =1),
                       plotto_signature_scoring_plot("Effector", merged3, size =1)))
```

```{r}
exhausted_int = list(pos = c("PDCD1","CTLA4","LAG3","CD160"),
                     minus = c("TCF7", "HAVCR2"))
merged3 = AddModuleScore(merged3, features = exhausted_int)
names(merged3@meta.data)[c(length(names(merged3@meta.data)) -1, length(names(merged3@meta.data)))] = c("Exhausted_int_pos", "Exhausted_int_neg")
merged3$Exhausted_int = scale(merged3$Exhausted_int_pos) - scale(merged3$Exhausted_int_neg)
ggarrange(plots = list(plotto_signature_scoring_plot("Exhausted_int_pos", merged3, size = 1),
                       plotto_signature_scoring_plot("Exhausted_int_neg", merged3, size =1),
                       plotto_signature_scoring_plot("Exhausted_int", merged3, size =1)))
```
```{r}
exhausted = list(plus = c("PDCD1", "HAVCR2", "LAG3", "TIGIT"),
                 minus = c("SELL", "TCF7", "CCR7", "IL17RA", "S1PR1",  "KLF2"))

merged3 = AddModuleScore(merged3, features = exhausted)
names(merged3@meta.data)[c(length(names(merged3@meta.data)) -1, length(names(merged3@meta.data)))] = c("Exhausted_pos", "Exhausted_neg")
merged3$Exhausted = scale(merged3$Exhausted_pos) - scale(merged3$Exhausted_neg)
ggarrange(plots = list(plotto_signature_scoring_plot("Exhausted_pos", merged3, size = 1),
                       plotto_signature_scoring_plot("Exhausted_neg", merged3, size =1),
                       plotto_signature_scoring_plot("Exhausted", merged3, size =1)))
```


```{r}
ifn_stim = list(plus = c("BST2", "IRF1", "IRF2", "IRF7", "STAT1", "STAT2", "MX1"))
merged3 = AddModuleScore(merged3, features = ifn_stim)
names(merged3@meta.data)[length(names(merged3@meta.data))] = "IFN_stimulated"

plotto_signature_scoring_plot("IFN_stimulated", merged3, size =1)
```
```{r}
trm = c("ITGAE","ITGA1", "CXCR6")
merged3 = AddModuleScore(merged3, features = list(trm))
names(merged3@meta.data)[length(names(merged3@meta.data))] = "TRM"
plotto_signature_scoring_plot("TRM", merged3, size =1)
```


```{r}
ggarrange(plots = list(plotto_marker_plot("TCF7", merged3, size = 1), plotto_marker_plot("HAVCR2", merged3, size = 1)))

```


```{r}
VlnPlot(merged3, features = c("Naive", "IFN_stimulated", "Effector", "Exhausted"), group.by = "condition", ncol = 2)
VlnPlot(merged3, features = c("Naive", "IFN_stimulated", "Effector", "Exhausted"),  ncol = 2)

RidgePlot(merged3, features = c("Naive", "IFN_stimulated", "Effector", "Exhausted"), sort = T, ncol = 2)
```

```{r}
# merged3$celltype = recode(merged3$seurat_clusters, 
#                           "7" = "naive", "9" = "naive", "3" = "naive", "8" = "naive",
#                           "0" = "effector", "2" = "effector", "6" = "effector", "10" = "effector",
#                           "5" = "exhausted")

merged3$celltype = recode(merged3$seurat_clusters, 
                          "7" = "TCM", "9" = "PT_Naive?", "3" = "TCM", "8" = "PT_Naive_activted?",
                          "0" = "Effector", "2" = "Effector", "6" = "Effector", "10" = "Effector",
                          "4" = "TEM",
                          "5" = "Exhausted")

DimPlot(merged3, group.by = "celltype", label = T)
```


### Harmony Integration
```{r}
library(harmony)
merged4 = RunHarmony(merged3, group.by.vars = "sample", dims.use = 1:30, plot_convergence = T, assay.use = "SCT")
merged4 = FindNeighbors(merged4, reduction = "harmony", dims = 1:25)
merged4 = RunUMAP(merged4, reduction = "harmony", dims = 1:25)

DimPlot(merged4, group.by = "condition", label = T)
```
### Seurat Integration
```{r}
split = Seurat::SplitObject(merged3, split.by = "sample")

split <- lapply(X = split, FUN = function(x) {
    x <- NormalizeData(x)
    x <- FindVariableFeatures(x, selection.method = "vst", nfeatures = 2000)
})

# select features that are repeatedly variable across datasets for integration
features <- SelectIntegrationFeatures(object.list = split)
features = features[!grepl("TR[AB]", features)]

immune.anchors <- FindIntegrationAnchors(object.list = split, anchor.features = features)
# this command creates an 'integrated' data assay
integrated_cd8s <- IntegrateData(anchorset = immune.anchors)

DefaultAssay(integrated_cd8s) = "integrated"
integrated_cd8s = ScaleData(integrated_cd8s)
integrated_cd8s = RunPCA(integrated_cd8s)
integrated_cd8s = RunUMAP(integrated_cd8s, dims = 1:25)

DimPlot(integrated_cd8s, group.by = "condition", label = T)
```


### Only PBMC
```{r}
merged4 = merged[, merged$condition == "PBMC" ]#& merged$sample != "PBMC B"& merged$sample != "PBMC A"]
merged4 = SCTransform(merged4)
varfeatures = merged4@assays$SCT@var.features[!grepl("TR[AB]", merged4@assays$SCT@var.features)]
merged4 = RunPCA(merged4, features = varfeatures)
merged4 = RunUMAP(merged4, dims = 1:20)

DimPlot(merged4, group.by = c("sample"), label = T)

FeaturePlot(merged4, features = c("CD8A", "CD4", "CD27", "CD14"))
```

```{r}
merged$expected_HTO_global = recode(merged$sample,
                                    "Donor 38" = "HTO-1", "Donor 41" = "HTO-2", "Donor 48" = "HTO-3", "Donor 49" = "HTO-4",
                                    "Donor 54" = "HTO-1_HTO-2", "Donor 56" = "HTO-2_HTO-3", "Donor 62" = "HTO-3_HTO-4",
                                    "Donor 64" = "HTO-1_HTO-4", "Donor 57" = "HTO-2_HTO-4") 
merged$expected_HTO_global[merged$orig.ident == "Sample_4A" & merged$sample == "Donor 41" & merged$HTO_classification == "HTO-1"] = "HTO-1"


subset = grepl("HTO",merged$expected_HTO_global)

summary(merged$HTO_classification[subset] == merged$expected_HTO_global[subset])

summary(merged$HTO_classification[subset] %in% c(merged$expected_HTO_global[subset], rep("Negative", 11426) ) )

conc = 0
disc = 0
for(c in rownames(merged@meta.data)[subset]){
  if(merged$HTO_classification[c] %in% c(merged$expected_HTO_global[c], "Negative", strsplit(merged$expected_HTO_global[c], "_")[[1]]))
    conc = conc + 1
  else
    disc = disc + 1
}
```


```{r}
merged$expected_HTO_local = recode(merged$Donor_ID,
                                    "Donor_38" = "HTO-1", "Donor_41" = "HTO-2", "Donor_48" = "HTO-3", "Donor_49" = "HTO-4",
                                    "Donor_54" = "HTO-1_HTO-2", "Donor_56" = "HTO-2_HTO-3", "Donor_62" = "HTO-3_HTO-4",
                                    "Donor_64" = "HTO-1_HTO-4", "Donor_57" = "HTO-2_HTO-4") 
merged$expected_HTO_local[merged$orig.ident == "Sample_4A" & merged$Donor_ID == "Donor_41" & merged$HTO_classification == "HTO-1"] = "HTO-1"


subset = grepl("HTO",merged$expected_HTO_local)


conc2 = 0
disc2 = 0
for(c in rownames(merged@meta.data)[subset]){
  if(merged$HTO_classification[c] %in% c(merged$expected_HTO_local[c], "Negative", strsplit(merged$expected_HTO_local[c], "_")[[1]]))
    conc2 = conc2 + 1
  else
    disc2 = disc2 + 1
}
```

## Indiviudal Donors

```{r}
donors = lapply(as.list(levels(factor(merged3$sample))), function(x){ 
  Don = merged3[,merged3$sample == x]
  Don = SCTransform(Don)
  varfeatures = Don@assays$SCT@var.features[!grepl("TR[AB]", Don@assays$SCT@var.features)]
  Don = RunPCA(Don, verbose = F, features = varfeatures)
  Don = RunUMAP(Don, dims=1:15)
  Don
})
```


```{r}
donors[[7]] = FindNeighbors(donors[[7]], dims=1:15)
donors[[7]] = FindClusters(donors[[7]])
DimPlot(donors[[7]])
m7 = FindAllMarkers(donors[[7]], only.pos = T)
m7 %>% group_by(cluster) %>% slice_head(n=5)
FeaturePlot(donors[[7]], features = c("KLRC2", "LTB","CD8A", "SRGN")) & scale_color_gradientn(colours = brewer.pal(9, "YlGn"))
```
```{r}
saveRDS(donors, "../output/donors.Rds")
```


## Pseudo Bulk

```{r}
library(edgeR)
s = levels(factor(paste(merged2$orig.ident, merged2$sample, merged2$condition)))

bulks = lapply(as.list(s), function(x) {
  b = rowSums(as.matrix(merged2@assays$RNA@counts[,paste(merged2$orig.ident, merged2$sample, merged2$condition) == x]))
})
names(bulks) = s

b = reduce(bulks, cbind)
colnames(b) = s

dge = DGEList(b)

dge$samples$group = factor(gsub("Sample_[1-4][AB] ","",rownames(dge$samples)) %>% gsub(" PBMC$","",.))
dge$samples$donor = factor(ifelse(grepl("Donor", dge$samples$group), substr(dge$samples$group, 1, 8), as.character(dge$samples$group)))
dge$samples$condition = factor(gsub("Donor [0-9]* ","", dge$samples$group) %>% gsub(" [A-G]$","", .))
dge$samples$run = factor(substr(rownames(dge$samples), 1, 9))

dge = dge[filterByExpr(dge),]
dge = calcNormFactors(dge)
```

```{r}
cpm = cpm(dge, log=T)
cpm = cpm[!grepl("TR[AB]", rownames(cpm)),]
pca = prcomp(t(cpm))

pca$x %>% as_tibble(rownames = "sample") %>%
  mutate(donor = dge$samples$donor, condition = dge$samples$condition) %>%
  ggplot(aes(PC1, PC2)) +
  geom_point(aes(color = donor))+
  theme_classic()
```
```{r}
# We filter PBMC B because it's so different.
# We further filter PBMC A because it too is not well aligned with the other PBMC samples 
# (it has additional cell types not conatined in the other, such as non CD8s)
# We also filter C01D-1 because it has no replicates in the data.
# Finally we filter two samples that have low cell counts.
dge2 = dge[,dge$samples$donor != "PBMC B" & dge$samples$condition != "C01D-1" & 
             dge$samples$donor != "PBMC A"]
dge2 = dge2[,-c(28,32,48)]
dge2 = dge2[filterByExpr(dge2),]
dge2 = calcNormFactors(dge2)
```

```{r}
cpm2 = cpm(dge2, log=T)
cpm2 = cpm2[!grepl("TR[AB]", rownames(cpm2)),]
pca2 = prcomp(t(cpm2))

pca2$x %>% as_tibble(rownames = "sample") %>%
  mutate(donor = dge2$samples$donor, condition = dge2$samples$condition) %>%
  group_by(condition) %>%
  summarise(PC1=mean(PC1), PC2=mean(PC2)) -> conlab

pca2$x %>% as_tibble(rownames = "sample") %>%
  mutate(donor = dge2$samples$donor, condition = dge2$samples$condition) %>%
  ggplot(aes(PC1, PC2)) +
  geom_point(aes(color = condition))+
  geom_text_repel(data=conlab, aes(label = condition, colour = condition), show.legend = F) +
  theme_classic() -> p1

pca2$x %>% as_tibble(rownames = "sample") %>%
  mutate(donor = dge2$samples$donor, condition = dge2$samples$condition) %>%
  group_by(donor) %>%
  summarise(PC1=mean(PC1), PC2=mean(PC2)) -> donlab

pca2$x %>% as_tibble(rownames = "sample") %>%
  mutate(donor = dge2$samples$donor, condition = dge2$samples$condition) %>%
  ggplot(aes(PC1, PC2)) +
  geom_point(aes(color = donor))+
  geom_text_repel(data=donlab, aes(label = donor, colour = donor), show.legend = F) +
  theme_classic() -> p2

p2+ p1
```

```{r}
library(sva)
dge2$samples$donor = factor(as.character(dge2$samples$donor))
dge2$samples$condition = factor(as.character(dge2$samples$condition))
normcpm = removeBatchEffect(cpm2, batch = dge2$samples$donor, batch2 = dge2$samples$run, design = model.matrix(~0+dge2$samples$condition))
normcpm = ComBat(cpm(dge2, log = T), batch = dge2$samples$donor, mod = model.matrix(~dge2$samples$condition))

pca3 = prcomp(t(normcpm))

pca3$x %>% as_tibble(rownames = "sample") %>%
  mutate(donor = dge2$samples$donor, condition = dge2$samples$condition) %>%
  group_by(condition) %>%
  summarise(PC1=mean(PC1), PC2=mean(PC2)) -> conlab

pca3$x %>% as_tibble(rownames = "sample") %>%
  mutate(donor = dge2$samples$donor, condition = dge2$samples$condition) %>%
  ggplot(aes(PC1, PC2)) +
  geom_point(aes(color = condition))+
  geom_text_repel(data=conlab, aes(label = condition, colour = condition), show.legend = F) +
  theme_classic() -> p1

pca3$x %>% as_tibble(rownames = "sample") %>%
  mutate(donor = dge2$samples$donor, condition = dge2$samples$condition) %>%
  group_by(donor) %>%
  summarise(PC1=mean(PC1), PC2=mean(PC2)) -> donlab

pca3$x %>% as_tibble(rownames = "sample") %>%
  mutate(donor = dge2$samples$donor, condition = dge2$samples$condition) %>%
  ggplot(aes(PC1, PC2)) +
  geom_point(aes(color = donor))+
  geom_text_repel(data=donlab, aes(label = donor, colour = donor), show.legend = F) +
  theme_classic() -> p2

p2 + p1
```

```{r}
library(edgeR)
s = levels(factor(paste(merged3$orig.ident, merged3$sample, merged3$condition, merged3$celltype)))

bulks = lapply(as.list(s), function(x) {
  b = rowSums(as.matrix(merged3@assays$RNA@counts[,paste(merged3$orig.ident, merged3$sample, merged3$condition, merged3$celltype) == x]))
})
names(bulks) = s

b = reduce(bulks, cbind)
colnames(b) = s

dge = DGEList(b)

dge$samples$group = factor(gsub("Sample_[1-4][AB] ","",rownames(dge$samples)))
dge$samples$donor = factor(ifelse(grepl("Donor", dge$samples$group), substr(dge$samples$group, 1, 8), as.character(dge$samples$group)))
dge$samples$condition = factor(gsub("Donor [0-9]* ","", dge$samples$group) %>% gsub(" [A-G]$","", .))
dge$samples$celltype = gsub("C[0-9]*D[+-][0-9]* ","", dge$samples$condition) %>% gsub("Baseline ","",.)
dge$samples$condition = gsub(" [0-9,A-Z,a-z,?,_]*$","", dge$samples$condition)
dge$samples$run = factor(substr(rownames(dge$samples), 1, 9))
dge$samples$group = factor(gsub("Donor [0-9]* ","",as.character(dge$samples$group)))

dge = dge[, dge$samples$lib.size > 4000]

dge = dge[filterByExpr(dge),]
dge = calcNormFactors(dge)
```

```{r}
library(sva)

normcpm = removeBatchEffect(cpm(dge, log = T), batch = dge$samples$donor, batch2 = dge$samples$run, design = model.matrix(~0+dge$samples$condition+dge$samples$celltype))
normcpm = ComBat(cpm(dge, log = T), batch = dge$samples$donor, mod = model.matrix(~dge$samples$condition+dge$samples$celltype))

# pca3 = prcomp(t(cpm(dge, log = T)))
pca3 = prcomp(t(normcpm[,dge$samples$celltype == "TCM"]))

pca3$x %>% as_tibble(rownames = "sample") %>%
  mutate(condition = dge$samples$condition[dge$samples$celltype == "TCM"]) %>%
  group_by(condition) %>%
  summarise(PC1=mean(PC1), PC2=mean(PC2)) -> conlab

pca3$x %>% as_tibble(rownames = "sample") %>%
  mutate(condition = dge$samples$condition[dge$samples$celltype == "TCM"]) %>%
  ggplot(aes(PC1, PC2)) +
  geom_point(aes(color = condition))+
  geom_text_repel(data=conlab, aes(label = condition, colour = condition), show.legend = F) +
  theme_classic() -> p1

pca3$x %>% as_tibble(rownames = "sample") %>%
  mutate(donor = dge$samples$donor, condition = dge$samples$condition) %>%
  group_by(donor) %>%
  summarise(PC1=mean(PC1), PC2=mean(PC2)) -> donlab

pca3$x %>% as_tibble(rownames = "sample") %>%
  mutate(donor = dge$samples$donor, condition = dge$samples$condition) %>%
  ggplot(aes(PC1, PC2)) +
  geom_point(aes(color = donor))+
  geom_text_repel(data=donlab, aes(label = donor, colour = donor), show.legend = F) +
  theme_classic() -> p2


pca3$x %>% as_tibble(rownames = "sample") %>%
  mutate(donor = dge$samples$donor, celltype = dge$samples$celltype) %>%
  group_by(celltype) %>%
  summarise(PC1=mean(PC1), PC2=mean(PC2)) -> ctlab

pca3$x %>% as_tibble(rownames = "sample") %>%
  mutate(donor = dge$samples$donor, celltype = dge$samples$celltype) %>%
  ggplot(aes(PC1, PC2)) +
  geom_point(aes(color = celltype))+
  geom_text_repel(data=ctlab, aes(label = celltype, colour = celltype), show.legend = F) +
  theme_classic() -> p3

p2 + p1 + p3
```


```{r}
G = factor(paste(dge$samples$celltype, dge$samples$condition, sep = "_") %>% 
             gsub("[?]","",.) %>% gsub("[+]","p",.) %>% gsub("[-]","m",.))
R = dge$samples$run
D = gsub(" ","_",dge$samples$donor)
design = model.matrix(~0+G+R + D)

v = voom(dge, design)

fit = lmFit(v, design)
cfit = contrasts.fit(fit, contrasts = makeContrasts(
  "C01+1" = GEffector_C01Dp1 - GEffector_Baseline,
  "C01+8" = GEffector_C01Dp8 - GEffector_Baseline,
  "C04-1" = GEffector_C04Dm1 - GEffector_Baseline,
  "C04+1" = GEffector_C04Dp1 - GEffector_Baseline,
  "C13-1" = GEffector_C13Dm1 - GEffector_Baseline,
  "C13+1" = GEffector_C13Dp1 - GEffector_Baseline,
  "C22+1" = GEffector_C22Dp1 - GEffector_Baseline,
  levels = design
))
efit = eBayes(cfit)

summary(decideTests(efit, p.value = 0.5))
```
```{r}
fit = lmFit(v, design)
cfit = contrasts.fit(fit, contrasts = makeContrasts(
  "C01+1" = GTCM_C01Dp1 - GTCM_Baseline,
  "C01+8" = GTCM_C01Dp8 - GTCM_Baseline,
  "C04-1" = GTCM_C04Dm1 - GTCM_Baseline,
  "C04+1" = GTCM_C04Dp1 - GTCM_Baseline,
  "C13-1" = GTCM_C13Dm1 - GTCM_Baseline,
  "C13+1" = GTCM_C13Dp1 - GTCM_Baseline,
  "C22+1" = GTCM_C22Dp1 - GTCM_Baseline,
  levels = design
))
efit = eBayes(cfit)

summary(decideTests(efit, p.value = 0.5))
```


### RUV III
Does not work with this replicate structure :(

```{r}
library(ruv)


dge2$samples$condition2 = factor(gsub("[+]","_plus_",dge2$samples$condition) %>% gsub("[-]","_minus_",.))

r = replicate.matrix(dge2$samples$condition2)
rownames(r) = rownames(dge2$samples)
ctl =  head(row.names(cpm(dge2)), n = 1000)

ruv_3 = RUVIII(Y = t(cpm(dge2, log = T)), M = r, ctl = ctl, k = 1, inputcheck = T,)
```

## RUV iv

```{r}
ruv_4 = RUV4(t(cpm(dge2, log = T)), X = r, ctl = ctl, k = 1)
```

## DGE Analysis

```{r}
G = dge2$samples$condition
R = dge2$samples$run
D = dge2$samples$donor

design = model.matrix(~G+D+R)

v = voom(dge2, design, plot = T)

fit = lmFit(v, design)

fit = contrasts.fit(fit, coefficients = c(2:8))

efit = eBayes(fit)

summary(decideTests(efit, p.value = 0.5))


topTable(efit, coef = 1)
```
```{r}
library(org.Hs.eg.db)
efit$entrez = mapIds(org.Hs.eg.db, rownames(efit), "SYMBOL", column = "ENTREZID")
kegg = kegga(efit, coef = 1, FDR = 0.5, species = "Hs", geneid = efit$entrez)
topKEGG(kegg)
```

### No PBMC

```{r}
dge3 = dge2[,dge2$samples$condition != "PBMC"]
G = factor(dge3$samples$condition)
R = factor(dge3$samples$run)
D = factor(dge3$samples$donor)

design = model.matrix(~G+D+R)

v = voom(dge3, design, plot = T)

fit = lmFit(v, design)

fit = contrasts.fit(fit, coefficients = c(2:21))

efit = eBayes(fit)

summary(decideTests(efit, p.value = 0.5))


topTable(efit, coef = 1)
```

```{r}
plotgene = function(g) {
  x = data.frame(X = normcpm[g,], donor = dge2$samples$donor, condition = dge2$samples$condition)
  x$s = paste(x$donor, x$condition)
  ggplot(x, aes(donor, X)) + 
    stat_summary(fun = mean, geom="bar", aes(fill= condition)) +
    geom_point() + 
    facet_wrap(~condition, scales = "free_x") + 
    theme_classic() +
    theme(axis.text.x = element_text(angle = 45, hjust = 1),
          legend.position = "none") +
    labs(x = element_blank(), y = paste0("log2 CPM ",g))
}

plotgene2 = function(g) {
  x = data.frame(X = normcpm[g,], donor = dge2$samples$donor, condition = dge2$samples$condition)
  ggplot(x, aes(condition, X)) + 
    stat_summary(fun = mean, geom="bar", aes(fill= condition)) +
    geom_point() + 
    theme_classic() +
    theme(axis.text.x = element_text(angle = 45, hjust = 1),
          legend.position = "none") +
    labs(x = element_blank(), y = paste0("log2 CPM ",g))
}
plotgene2("CARD16")
```
PBMCs by themselves...
CRISPR Dead Cas 9 off target?

```{r}
FindMarkers(merged3, ident.1 = "C01D+1", ident.2 = "Baseline", group.by = "condition", latent.vars = c("sample", "orig.ident"), test.use = "LR") -> m

m2 = topTable(efit, coef = 1, number = 117)

vennDiagram(data.frame(pseudo_bulk = ifelse(row.names(efit) %in% rownames(m2), TRUE, FALSE), 
                       single_cell = ifelse(row.names(efit) %in% rownames(m), TRUE, FALSE)))
```


## Functional readouts
```{r}
func = read.csv("../data/2022-07-18_AMC_ICS_Cleaned_Readout.csv")
func = func[-26,]

pca = prcomp(func[,-c(1:5)])

pca$x %>% as_tibble() %>%
  mutate(donor = func$Donor, condition = func$Timepoint) %>%
  ggplot(aes(PC1, PC2)) +
  geom_point(aes(colour = condition)) +
  theme_classic()
```

pca$rotation = 
                  PC1          PC2          PC3          PC4         PC5         PC6
CD4_IFNg -0.027042357  0.002431816 -0.005844169 -0.438726324 -0.28160233  0.85290561
CD4_TNFa -0.032694258 -0.015479651 -0.010315274 -0.799827739 -0.30759906 -0.51404617
CD4_IL2   0.005408232  0.689273840 -0.141498643 -0.296038446  0.64336588  0.05737633
CD8_IFNg -0.731195476 -0.212360244 -0.641474812  0.002893525  0.09343038  0.00536280
CD8_TNFa -0.679832658  0.198664680  0.704109543  0.035260746 -0.03508457 -0.01074273
CD8_IL2  -0.036759755  0.663393934 -0.269401915  0.280899175 -0.63419487 -0.06980215

CD8_IFNg and CD8_TNFa drive the PC1 variance.
CD4_IL2 and CD8_IL2 for PC2

```{r}
func[,-c(3:5)] %>%
  pivot_longer(-c(Donor, Timepoint)) %>%
  ggplot(aes(Timepoint, value)) +
  geom_point(aes(colour = Donor)) +
  geom_line(aes(group = Donor, colour = Donor)) +
  facet_wrap(~name) +
  theme_classic() +
  theme(axis.text.x = element_text(angle = 45, hjust = 1))
  
```



```{r}
func$group = paste(func$Donor, func$Timepoint) %>% gsub("Donor", "Donor ",. )
func %>% group_by(group) %>%
  summarise_all(mean) -> func2

dge_bulk$samples %>% as_tibble() %>% left_join(func2, by = "group") -> x

dge_func = dge_bulk

dge_func$samples = cbind(dge_func$samples, x[,9:14])

dge_func = dge_func[,!is.na(dge_func$samples$CD8_IL2)]

normcpm = removeBatchEffect(cpm(dge_func, log = T), batch = as.numeric(dge_func$samples$CD8_IFNg), batch2 = as.numeric(dge_func$samples$CD8_IL2), design = model.matrix(~dge_func$samples$group))

G = factor(dge_func$samples$condition)
R = factor(dge_func$samples$run)
D = factor(dge_func$samples$donor)

design = model.matrix(~G+D+R+dge_func$samples$CD8_IFNg+dge_func$samples$CD8_IL2+dge_func$samples$CD4_IFNg+ dge_func$samples$CD4_TNFa+dge_func$samples$CD4_IL2+dge_func$samples$CD8_TNFa)

v = voom(dge_func, design, plot = T)

fit = lmFit(v, design)

fit = contrasts.fit(fit, coefficients = c(2:21))

efit = eBayes(fit)

summary(decideTests(efit, p.value = 0.05))
```

## Celltype Proportions

```{r}
merged3@meta.data %>% as_tibble() %>%
  group_by(condition) %>% mutate(norm = 1/n()) %>% ungroup() %>%
  group_by(celltype) %>% mutate(tot = sum(norm)) %>%
  group_by(celltype, condition) %>% summarise(f = sum(norm)/tot[1]) %>%
  ggplot(aes(celltype, f)) +
  geom_bar(stat="identity", aes(fill = condition)) +
  theme_classic()
  

merged3@meta.data %>% as_tibble() %>%
  group_by(condition) %>% mutate(norm = 1/n()) %>% ungroup() %>%
  group_by(celltype) %>% mutate(tot = sum(norm)) %>%
  group_by(celltype, condition) %>% summarise(f = sum(norm)/tot[1]) %>%
  ggplot(aes(condition,f)) +
  geom_point(aes(colour = celltype)) +
  geom_line(aes(colour = celltype, group = celltype)) +
  facet_wrap(~celltype, ncol = 5)+
  theme_classic() +
  theme(axis.text.x = element_text(angle = 45, hjust = 1), legend.position = "none")
```


```{r}
effectors = merged3[,merged3$celltype == "Effector"]
effectors@active.assay = "RNA"
effectors %>% NormalizeData() %>% FindVariableFeatures() -> effectors
varfeatures = effectors@assays$RNA@var.features[!grepl("TR[AB]", effectors@assays$RNA@var.features)]

effectors = ScaleData(effectors) %>% RunPCA( features = varfeatures, verbose = F) %>% RunUMAP(dims = 1:20) 

DimPlot(effectors, group.by = c("condition", "sample"), label = T)

library(harmony)
effectors = RunHarmony(effectors, group.by.vars = "sample")
effectors = RunUMAP(effectors, reduction = "harmony", dims = 1:20)
```

## CytoTRACE

```{r}
library(CytoTRACE)

CytoTRACE = CytoTRACE(as.matrix(merged3@assays$SCT@data))

merged3$CytoTRACE = CytoTRACE$CytoTRACE
plotto_signature_scoring_plot("CytoTRACE", merged3, size = 1)
```


## Shiny App
```{r}
library(ShinyCell)

scConf = createConfig(merged3)

makeShinyApp(merged3, scConf, gene.mapping = TRUE,
             shiny.title = "HIV Cancer CD8 T Cells") 
```

