---
title: "06_reservoir_data"
author: "jibsch"
date: "2023-01-31"
output: workflowr::wflow_html
editor_options:
  chunk_output_type: inline
---

## Introduction

```{r}
library(tidyverse)
library(openxlsx)
library(ggrepel)
library(Seurat)
library(edgeR)

merged4 <- readRDS("/data/gpfs/projects/punim1612/Lewinlab/Lewinlab_AMC_local/output/merged4.Rds")
RNA_dat_summarised2 = readRDS("../output/RNA_dat_summarised2.Rds")
DNA_dat_summarised4 = readRDS("../output/DNA_dat_summarised4.Rds")

donor_mapping = c("Donor 38" = "P1", "Donor 41" = "P2", "Donor 49" = "P4",
                  "Donor 56" = "P6", "Donor 62" = "P7", "Donor 64" = "P8")

source("~/tools/BioinfTools/Plotting_Functions.R")
```
Here we plot the functional data analysis values for responders and non-responders
The data went through a few iterations... See the second round of ggsaves for the final 
plots (adding additional donors)


```{r}
data = readWorkbook("../data/AMC 095 HIV DNA_updated 19 Dec 2019.xlsx", startRow = 5)

seq = paste0("C",c(1,4,5,7,9,13,16,19,22,25,37,40))

names(data)[5:40] = paste(c(rbind(seq,seq,seq)), names(data)[5:40], sep = "_")

data %>% as_tibble() %>% filter(grepl("0(38|41|48|49|54|56|62|64)$", Study.ID)) %>% 
  pivot_longer(cols = starts_with("C") | starts_with("D") | starts_with("A")) %>%
  filter(!is.na(value)) %>%
  mutate(name = factor(name)) %>% mutate(name = factor(name, levels = levels(name)[c(1,9,2:4,6:8,5)])) %>%
  mutate(donor = paste("Donor", substr(Study.ID,6,8))) %>%
  mutate(response = ifelse(donor %in% c("Donor 49", "Donor 56", "Donor 38"), "Response", "No response")) %>%
  
  mutate(name = recode(name, 
                       "C1_D-1" = "D0.(BL)")) %>%
  
  filter(name != "C13_D-1") -> RNA_dat
  
ggplot(RNA_dat, aes(name, value)) + 
  geom_point() + geom_line(aes(group = donor)) + 
  facet_wrap(response~donor, scales = "free_y") +
  theme_classic()+
  theme(axis.text.x = element_text(angle = 60, hjust = 1))
```


```{r}
RNA_dat %>%
  mutate(time = factor(ifelse(name %in% c("Avg.pre-infusion" ,         "D0.(BL)"), "No dose",
                       ifelse(grepl("C1_", name), "Single dose",  "Multi dose")),
                       levels = c("No dose", "Single dose",  "Multi dose"))) %>%
  mutate(donor = factor(donor, levels = c("Donor 38", "Donor 49", "Donor 56",
                                            "Donor 41", "Donor 48", "Donor 54", "Donor 62", "Donor 64"))) %>%
  group_by(donor, time, response) %>%
  summarise(value = mean(value)) -> RNA_dat_summarised



ggplot(RNA_dat_summarised, aes(time, value)) +
  geom_bar(data = RNA_dat_summarised %>% group_by(time, response) %>% summarise(value = mean(value)),
           stat = "identity",
           aes(fill = time)) +
  geom_point(aes(colour = donor)) +
  geom_line(aes(group = donor, colour = donor)) +
  facet_wrap(~response) +
  scale_colour_manual(values = colour_donors) +
  scale_fill_manual(values = c("lightgrey", "grey", "darkgrey")) +
  scale_y_continuous(trans = "log10") +
  theme_classic() +
  theme(axis.text.x = element_text(angle = 45, hjust = 1)) ->p;p

ggsave(p, filename = "../plots/RNA_functional_BSM_log.pdf", width = 15, height = 9, units = "cm")
```


## RNA

```{r}
data2 = readWorkbook("../data/US RNA replicate data incl 18s_averaged estimates incl.xlsx", startRow = 5)

for(i in 1:length(data2$Study.ID)) {
  if(is.na(data2$Study.ID[i])) 
    data2[i,1] = last
  else 
    last = data2$Study.ID[i]
}

seq = paste0("C",c(1,4,5,7,9,13,16,19,22,25,37,40))

names(data)[5:40] = paste(c(rbind(seq,seq,seq)), names(data)[5:40], sep = "_")

data2 %>% 
  as_tibble() %>% filter(grepl("-0(38|41|48|49|54|56|62|64)$", Study.ID)) %>% 
  mutate(donor = paste("Donor", substr(Study.ID,10,11))) %>%
  mutate(response = factor(ifelse(donor %in% c("Donor 49", "Donor 56", "Donor 38"), "Response", ifelse( donor %in% c("Donor 48", "Donor 54"),"Unclassified","No response")),
                          levels = c("Response", "No response", "Unclassified"))) %>% 
  
  filter(!is.na(`average.US.HIV.copies/10^6.18S`)) %>% 
  
  select(donor, Timepoint, response, "average.US.HIV.copies/10^6.18S", "Average.number.US.RNA.copies") %>%
  
  mutate(Timepoint = recode(Timepoint, 
                       "C1D-1" = "BL")) %>%
  mutate(time = factor(ifelse(Timepoint %in% c("BL"), "No dose",
                       ifelse(grepl("C1", Timepoint), "Single dose",  "Multi dose")),
                       levels = c("No dose", "Single dose",  "Multi dose"))) -> RNA_dat
  
# ggplot(RNA_dat, aes(Timepoint, average.US.HIV.copies/10^6.18S)) + 
#   geom_point() + geom_line(aes(group = donor)) + 
#   facet_wrap(response~donor, scales = "free_y") +
#   theme_classic()+
#   theme(axis.text.x = element_text(angle = 60, hjust = 1))
```

```{r}
RNA_dat %>%
  mutate(donor = factor(donor, levels = c("Donor 38", "Donor 49", "Donor 56",
                                            "Donor 41", "Donor 62", "Donor 64", "Donor 48", "Donor 54"))) %>%
  group_by(donor, time, response) %>%
  summarise(value = mean(`average.US.HIV.copies/10^6.18S`), c = mean(Average.number.US.RNA.copies)) -> RNA_dat_summarised



ggplot(RNA_dat_summarised, aes(time, value)) +
  geom_bar(data = RNA_dat_summarised %>% group_by(time, response) %>% summarise(value = mean(value)),
           stat = "identity",
           aes(fill = time)) +
  geom_point(aes(colour = donor)) +
  geom_line(aes(group = donor, colour = donor)) +
  facet_wrap(~response) +
  scale_colour_manual(values = colour_donors) +
  scale_fill_manual(values = c("lightgrey", "grey", "darkgrey")) +
  # scale_y_continuous(trans = "log10") +
  theme_classic() +
  labs(y = "average.US.HIV.copies/10^6.18S") +
  theme(axis.text.x = element_text(angle = 45, hjust = 1)) ->p;p

ggsave(p, filename = "../plots/RNA_functional_BSM.pdf", width = 15, height = 9, units = "cm")
```

AMC095_062 and 064.xlsx not used - see below

```{r}
data5 = readWorkbook("../data/AMC095_062 and 064 updated qPCR data.xlsx", startRow = 4, sheet = 2)

data5 %>% 
  as_tibble() %>% filter(grepl("-0(38|41|48|49|54|56|62|64)$", patient)) %>% 
  mutate(donor = paste("Donor", substr(patient,10,11))) %>%
  mutate(response = factor(ifelse(donor %in% c("Donor 49", "Donor 56", "Donor 38"), "Response", 
                                  ifelse( donor %in% c("Donor 48", "Donor 54"),"Unclassified", "No response")),
                          levels = c("Response", "No response", "Unclassified"))) %>% 
  
  select(donor, visit, response, "average.US.HIV.RNA.copies./10^6.18S") %>%
  
  mutate(visit = recode(visit, 
                       "C1D-1" = "BL", "C1D-1/Baseline" = "BL")) %>%
  mutate(time = factor(ifelse(visit %in% c("BL"), "No dose",
                       ifelse(grepl("C1", visit), "Single dose",  "Multi dose")),
                       levels = c("No dose", "Single dose",  "Multi dose"))) -> RNA_dat2

RNA_dat %>% select(-"Average.number.US.RNA.copies", -"Timepoint") %>%
  rbind(RNA_dat2 %>% select(-"visit") %>% rename("average.US.HIV.copies/10^6.18S" = "average.US.HIV.RNA.copies./10^6.18S" )) %>% 
  mutate(donor = factor(donor, levels = c("Donor 38", "Donor 49", "Donor 56",
                                            "Donor 41",  "Donor 62", "Donor 64", "Donor 48", "Donor 54"))) %>%
  group_by(donor, time, response) %>%
  summarise(value = mean(`average.US.HIV.copies/10^6.18S`)) -> RNA_dat_summarised2


ggplot(RNA_dat_summarised2, aes(time, value)) +
  geom_bar(data = RNA_dat_summarised2 %>% group_by(time, response) %>% summarise(value = median(value)),
           stat = "identity",
           aes(fill = time)) +
  geom_point(aes(colour = donor)) +
  geom_line(aes(group = donor, colour = donor)) +
  facet_wrap(~response) +
  scale_colour_manual(values = colour_donors) +
  scale_fill_manual(values = c("lightgrey", "grey", "darkgrey")) +
  # scale_y_continuous(trans = "log10") +
  theme_classic() +
  labs(y = "average.US.HIV.copies/10^6.18S") +
  theme(axis.text.x = element_text(angle = 45, hjust = 1)) ->p;p

ggsave(p, filename = "../plots/RNA_functional_BSM.pdf", width = 18, height = 9, units = "cm")
```
```{r}
RNA_dat_summarised2 %>% 
  filter(!donor %in% c("Donor 48", "Donor 54")) %>%
  mutate(donor = donor_mapping[as.character(donor)]) %>%
  ggplot( aes(time, value)) +
  geom_bar(data = RNA_dat_summarised2 %>% group_by(time) %>% summarise(value = median(value)),
           stat = "identity",
           aes(fill = time)) +
  geom_point(aes(shape = donor)) +
  geom_line(aes(group = donor)) +
  scale_fill_manual(values = c("lightgrey", "grey", "darkgrey")) +
  # scale_y_continuous(trans = "log10") +
  theme_classic() +
  labs(y = "average.US.HIV.copies/10^6.18S") +
  theme(axis.text.x = element_text(angle = 45, hjust = 1)) ->p;p

ggsave(p, filename = "../plots/RNA_functional_BSM_no_groups.pdf", width = 10, height = 9, units = "cm")
```

## DNA2

```{r}
data3 = readWorkbook("../data/HIV DNA replicate data incl CCR5_averaged estimates incl.xlsx", startRow = 3)

for(i in 1:length(data3$Study.ID)) {
  if(is.na(data3$Study.ID[i])) 
    data3[i,1] = last
  else 
    last = data3$Study.ID[i]
}


data3 %>% 
  as_tibble() %>% filter(grepl("-0(38|41|48|49|54|56|62|64)$", Study.ID)) %>% 
  mutate(donor = paste("Donor", substr(Study.ID,10,11))) %>%
  mutate(response = factor(ifelse(donor %in% c("Donor 49", "Donor 56", "Donor 38"), "Response", ifelse( donor %in% c("Donor 48", "Donor 54"),"Unclassified","No response")),
                           levels = c("Response", "No response", "Unclassified"))) %>% 
  
  filter(!is.na(`average.HIV.copies/10^6.CCR5`)) %>% 
  
  select(donor, Timepoint, response, "average.HIV.copies/10^6.CCR5", "Average.number.HIV.DNA.copies") %>%
  
  mutate(Timepoint = recode(Timepoint, 
                       "C1 D-1" = "BL", "C1D-1" = "BL")) %>%
  mutate(time = factor(ifelse(Timepoint %in% c("BL"), "No dose",
                       ifelse(grepl("C1", Timepoint), "Single dose",  "Multi dose")),
                       levels = c("No dose", "Single dose",  "Multi dose"))) -> DNA_dat
  
# ggplot(RNA_dat, aes(Timepoint, average.US.HIV.copies/10^6.18S)) + 
#   geom_point() + geom_line(aes(group = donor)) + 
#   facet_wrap(response~donor, scales = "free_y") +
#   theme_classic()+
#   theme(axis.text.x = element_text(angle = 60, hjust = 1))
```

```{r}
DNA_dat %>%
  mutate(donor = factor(donor, levels = c("Donor 38", "Donor 49", "Donor 56",
                                            "Donor 41", "Donor 62", "Donor 64", "Donor 48", "Donor 54"))) %>%
  group_by(donor, time, response) %>%
  summarise(value = mean(`average.HIV.copies/10^6.CCR5`), c = mean(Average.number.HIV.DNA.copies)) -> DNA_dat_summarised



ggplot(DNA_dat_summarised, aes(time, value)) +
  geom_bar(data = DNA_dat_summarised %>% group_by(time, response) %>% summarise(value = mean(value)),
           stat = "identity",
           aes(fill = time)) +
  geom_point(aes(colour = donor)) +
  geom_line(aes(group = donor, colour = donor)) +
  facet_wrap(~response) +
  scale_colour_manual(values = colour_donors) +
  scale_fill_manual(values = c("lightgrey", "grey", "darkgrey")) +
  # scale_y_continuous(trans = "log10") +
  theme_classic() +
  labs(y = "average.HIV.copies/10^6.CCR5") +
  theme(axis.text.x = element_text(angle = 45, hjust = 1)) ->p;p

ggsave(p, filename = "../plots/DNA_functional_BSM.pdf", width = 15, height = 9, units = "cm")
```

"AMC095_062 and 064.xlsx" had a calculation error...

```{r}
data4 = readWorkbook("../data/AMC095_062 and 064 updated qPCR data.xlsx", startRow = 6)

data4 %>% 
  as_tibble() %>% filter(grepl("-0(38|41|48|49|54|56|62|64)$", Sample.ID)) %>% 
  mutate(donor = paste("Donor", substr(Sample.ID,10,11))) %>%
  mutate(response = factor(ifelse(donor %in% c("Donor 49", "Donor 56", "Donor 38"), "Response", ifelse( donor %in% c("Donor 48", "Donor 54"),"Unclassified", "No response")),
                           levels = c("Response", "No response", "Unclassified"))) %>%  
  
  select(donor, Timepoint, response,  "average.HIV.copies/10^.CCR5") %>%
  
  mutate(Timepoint = recode(Timepoint, 
                       "C1D-1" = "BL", "C1D-1/Baseline" = "BL")) %>%
  mutate(time = factor(ifelse(Timepoint %in% c("BL"), "No dose",
                       ifelse(grepl("C1", Timepoint), "Single dose",  "Multi dose")),
                       levels = c("No dose", "Single dose",  "Multi dose"))) -> DNA_dat4


DNA_dat %>% select(-"Average.number.HIV.DNA.copies") %>%
  rbind(DNA_dat4  %>% rename("average.HIV.copies/10^6.CCR5" = "average.HIV.copies/10^.CCR5" )) %>%
  mutate(donor = factor(donor, levels = c("Donor 38", "Donor 49", "Donor 56",
                                            "Donor 41", "Donor 62", "Donor 64", "Donor 48", "Donor 54"))) %>%
  group_by(donor, time, response) %>%
  summarise(value = mean(`average.HIV.copies/10^6.CCR5`)) -> DNA_dat_summarised4

ggplot(DNA_dat_summarised4, aes(time, value)) +
  geom_bar(data = DNA_dat_summarised4 %>% group_by(time, response) %>% summarise(value = median(value)),
           stat = "identity",
           aes(fill = time)) +
  geom_point(aes(colour = donor)) +
  geom_line(aes(group = donor, colour = donor)) +
  facet_wrap(~response) +
  scale_colour_manual(values = colour_donors) +
  scale_fill_manual(values = c("lightgrey", "grey", "darkgrey")) +
  # scale_y_continuous(trans = "log10") +
  theme_classic() +
  labs(y = "average.HIV.copies/10^6.CCR5") +
  theme(axis.text.x = element_text(angle = 45, hjust = 1)) ->p;p

ggsave(p, filename = "../plots/DNA_functional_BSM.pdf", width = 18, height = 9, units = "cm")
```

## RNA/DNA

```{r}
RNA_dat_summarised2 %>% inner_join(DNA_dat_summarised4, by = c("donor", "time", "response")) %>% 
  mutate(ratio = value.x/value.y) %>% 
  filter(!is.infinite(ratio)) ->ratio_dat
  
ggplot(ratio_dat, aes(time, ratio)) +
  geom_bar(data = ratio_dat %>% group_by(time, response) %>% summarise(ratio = mean(ratio)),
           stat = "identity",
           aes(fill = time)) +
  geom_point(aes(colour = donor)) +
  geom_line(aes(group = donor, colour = donor)) +
  facet_wrap(~response) +
  scale_colour_manual(values = colour_donors) +
  scale_fill_manual(values = c("lightgrey", "grey", "darkgrey")) +
  # scale_y_continuous(trans = "log10") +
  theme_classic() +
  labs(y = "RNA copies / DNA copies") +
  theme(axis.text.x = element_text(angle = 45, hjust = 1)) ->p;p

ggsave(p, filename = "../plots/RNA_vs_DNA_functional_BSM.pdf", width = 15, height = 9, units = "cm")
```


## CD8 Tet+ 

```{r}
library(openxlsx)

data = readWorkbook("../data/CD8 Tet+.xlsx")

data[-1, ] %>% as_tibble() %>% 
  group_by(X1) %>%
  summarise(BL = mean(Baseline, X4, na.rm = T, trim = 0.5))

data = data[-1,]
cd8 = data.frame(P = data$X1, 
                 B = rowMeans(data[,3:4], na.rm = T),
                 S = rowMeans(cbind(as.numeric(data[,5]), as.numeric(data[,6]), as.numeric(data[,7])) , na.rm = T),
                 M = rowMeans(cbind(as.numeric(data[,8]), as.numeric(data[,9]), as.numeric(data[,10]),
                                    as.numeric(data[,11]), as.numeric(data[,12]), as.numeric(data[,15])) , na.rm = T)
                 )

library(speckle)

prop.trans <- asin(sqrt(props))

cd8_trans = cd8 %>% mutate(B = asin(sqrt(B)), S = asin(sqrt(S)), M = asin(sqrt(M)))

G = 

design = model.matrix()

fit <- lmFit(cd8_trans,design)
```

## Correlations

```{r}
library(Seurat)
merged4 <- readRDS("~/projects/Lewinlab_AMC_local/output/merged4.Rds")

RNA_dat %>% select(-"Average.number.US.RNA.copies", -time) %>% 
  mutate(Timepoint = recode(Timepoint,
                            "C1D1" = "C1D+1",
                            "C1D7" = "C1D8",
                            "C4D1" = "C4D+1")) %>%
  rbind(RNA_dat2  %>% select(-time) %>% dplyr::rename("average.US.HIV.copies/10^6.18S" = "average.US.HIV.RNA.copies./10^6.18S" , "Timepoint" = "visit")) %>% 
  mutate(donor = factor(donor, levels = c("Donor 38", "Donor 49", "Donor 56",
                                            "Donor 41",  "Donor 62", "Donor 64", "Donor 48", "Donor 54"))) %>% 
  group_by(donor, Timepoint, response) %>%
  summarise(value = mean(`average.US.HIV.copies/10^6.18S`)) -> RNA_dat_summarised2

RNA_dat %>% select(-"Average.number.US.RNA.copies", -Timepoint) %>% 
  rbind(RNA_dat2  %>% select(-visit) %>% dplyr::rename("average.US.HIV.copies/10^6.18S" = "average.US.HIV.RNA.copies./10^6.18S" )) %>% 
  mutate(donor = factor(donor, levels = c("Donor 38", "Donor 49", "Donor 56",
                                            "Donor 41",  "Donor 62", "Donor 64", "Donor 48", "Donor 54"))) %>% 
  group_by(donor, time, response) %>%
  summarise(value = mean(`average.US.HIV.copies/10^6.18S`)) -> RNA_dat_summarised_bsm2
```


```{r}
merged4$Cluster1 %>% as_tibble() %>%
  mutate(donor = merged4$sample, Timepoint = merged4$Timepoint) %>% 
  mutate(Timepoint = recode(Timepoint,
                       "Baseline" = "BL",  "C01D-1" = "BL",
                       "C01D+1" = "C1D+1", "C01D+8" = "C1D8",
                       "C04D-1" = "C4D-1", "C04D+1" = "C4D+1")) %>%
  group_by(donor, Timepoint) %>% summarise(ISG = mean(value))  %>%
  inner_join(RNA_dat_summarised2) %>% 
  filter(response != "Unclassified") %>%
  ggplot(aes(ISG, value)) + geom_point(aes(color= Timepoint)) + facet_wrap(~response) + geom_line(aes(group = Timepoint, colour = Timepoint)) +
  geom_line(aes(group = donor), linetype = "dashed")

# No Time
merged4$Cluster1 %>% as_tibble() %>%
  mutate(donor = merged4$sample, Timepoint = merged4$Timepoint) %>% 
  mutate(Timepoint = recode(Timepoint,
                       "Baseline" = "BL",  "C01D-1" = "BL",
                       "C01D+1" = "C1D+1", "C01D+8" = "C1D8",
                       "C04D-1" = "C4D-1", "C04D+1" = "C4D+1")) %>%
  group_by(donor, Timepoint) %>% summarise(ISG = mean(value))  %>%
  inner_join(RNA_dat_summarised2) %>% 
  filter(response != "Unclassified") %>%
  ggplot(aes(ISG, value)) + geom_point() + facet_wrap(~response)  +
  geom_smooth()

merged4$Cluster1 %>% as_tibble() %>%
  mutate(donor = merged4$sample, Timepoint = merged4$Timepoint) %>% 
  mutate(Timepoint = recode(Timepoint,
                       "Baseline" = "BL",  "C01D-1" = "BL",
                       "C01D+1" = "C1D+1", "C01D+8" = "C1D8",
                       "C04D-1" = "C4D-1", "C04D+1" = "C4D+1")) %>%
  group_by(donor, Timepoint) %>% summarise(ISG = mean(value))  %>%
  inner_join(RNA_dat_summarised2) %>% 
  filter(response != "Unclassified") %>%
  group_by(response) %>%
  summarise(c = cor(ISG, value))

#MX1 gene
merged4@assays$SCT@data["MX1",] %>% as_tibble() %>%
  mutate(donor = merged4$sample, Timepoint = merged4$Timepoint) %>% 
  mutate(Timepoint = recode(Timepoint,
                       "Baseline" = "BL",  "C01D-1" = "BL",
                       "C01D+1" = "C1D+1", "C01D+8" = "C1D8",
                       "C04D-1" = "C4D-1", "C04D+1" = "C4D+1")) %>%
  group_by(donor, Timepoint) %>% summarise(MX1 = mean(value))  %>%
  inner_join(RNA_dat_summarised2) %>% 
  filter(response != "Unclassified") %>%
  ggplot(aes(MX1, value)) + geom_point(aes(color= Timepoint)) + facet_wrap(~response) + geom_line(aes(group = Timepoint, colour = Timepoint)) +
  geom_line(aes(group = donor), linetype = "dashed")

merged4@assays$SCT@data["STAT1",] %>% as_tibble() %>%
  mutate(donor = merged4$sample, time = merged4$time) %>% 
  group_by(donor, time) %>% summarise(MX1 = mean(value))  %>% 
  inner_join(RNA_dat_summarised_bsm2) %>%
  filter(response != "Unclassified") %>%
  group_by(response, time) %>%
  mutate(c = cor(MX1, value)) %>%
  ggplot(aes(MX1, value)) +  facet_wrap(~response) + 
  geom_line(aes(group = time, colour = time), size = 2) +
  geom_line(aes(group = donor), linetype = "dashed") +
  geom_point(aes(fill= donor), shape = 21, size = 3) +
  scale_colour_manual(values = c("lightgrey", "grey", "darkgrey")) +
  scale_fill_manual(values = colour_donors) +
  theme_classic() +
  labs(x = "STAT1 expression", y = "RNA levels average.US.HIV.RNA.copies./10^6.18S")->p;p
ggsave(p, filename = "../plots/Functional_data_STAT1_RNA.pdf", width = 15, height = 9, units = "cm")

merged4@assays$SCT@data["MX1",] %>% as_tibble() %>%
  mutate(donor = merged4$sample, time = merged4$time) %>% 
  group_by(donor, time) %>% summarise(MX1 = mean(value))  %>% 
  inner_join(RNA_dat_summarised_bsm2) %>%
  filter(response != "Unclassified") %>%
  group_by(response, time) %>%
  mutate(c = cor(MX1, value)) %>%
  ggplot(aes(MX1, value)) +  facet_wrap(~response) + 
  geom_line(aes(group = time, colour = time), size = 2) +
  geom_line(aes(group = donor), linetype = "dashed") +
  geom_point(aes(fill= donor), shape = 21, size = 3) +
  scale_colour_manual(values = c("lightgrey", "grey", "darkgrey")) +
  scale_fill_manual(values = colour_donors) +
  theme_classic() +
  labs(x = "MX1 expression", y = "RNA levels average.US.HIV.RNA.copies./10^6.18S")->p;p
ggsave(p, filename = "../plots/Functional_data_MX1_RNA.pdf", width = 15, height = 9, units = "cm")

merged4@assays$SCT@data["NAV2",] %>% as_tibble() %>%
  mutate(donor = merged4$sample, time = merged4$time) %>% 
  group_by(donor, time) %>% summarise(NAV2 = mean(value))  %>% 
  inner_join(RNA_dat_summarised2) %>%
  filter(response != "Unclassified") %>%
  group_by(response, time) %>%
  mutate(c = cor(NAV2, value)) %>%
  ggplot(aes(NAV2, value)) + geom_point(aes(color= time)) + 
  facet_wrap(~response) + geom_line(aes(group = time, colour = time))
```

```{r}
merged4@assays$SCT@data[merged4@assays$SCT@var.features,] %>% as_tibble(rownames = "gene") %>% 
  filter(!grepl("TR", gene)) %>%
  pivot_longer(-gene) %>%
  inner_join(as_tibble(merged4@meta.data, rownames = "name") %>% select(name, time, sample)) %>%
  rename("donor" = "sample") %>%
  group_by(donor, time, gene) %>% summarise(exp = mean(value))  %>% 
  inner_join(RNA_dat_summarised2) %>%
  filter(response != "Unclassified") %>%
  group_by(response, time, gene) %>%
  summarise(c = cor(exp, value), exp = mean(exp), v = mean(value)) %>% arrange(-c) -> cordat

cordat %>% group_by(gene, time) %>% summarise(diff = abs(c[1] - c[2]), exp=max(exp)) %>%  
  group_by(gene) %>% 
    mutate(ratio = diff[2]/diff[1]) %>%
  arrange(-ratio)%>%
  filter(exp > 1)
```
```{r}
plot_gene = function(g) {
  merged4@assays$SCT@data[g,] %>% as_tibble() %>%
  mutate(donor = merged4$sample, time = merged4$time) %>% 
  group_by(donor, time) %>% summarise(exp = mean(value))  %>% 
  inner_join(RNA_dat_summarised2) %>%
  filter(response != "Unclassified") %>%
  group_by(response, time) %>%
  mutate(c = cor(exp, value)) %>%
  ggplot(aes(exp, value)) + geom_point(aes(color= time)) + facet_wrap(~response) + geom_line(aes(group = time, colour = time)) + ggtitle(g)
}

plot_gene("PRKX")
plot_gene("ACTB")
plot_gene("IFITM1")
plot_gene("ISG15")
```
## Interferon Signature
```{r}
isg = c("ADAR","APOL6","ARID5B","ARL4A","AUTS2","B2M","BANK1","BATF2","BPGM","BST2","BTG1","C1R","C1S","CASP1","CASP3","CASP4","CASP7","CASP8","CCL2","CCL5","CCL7","CD274","CD38","CD40","CD69","CD74","CD86","CDKN1A","CFB","CFH","CIITA","CMKLR1","CMPK2","CSF2RB","CXCL10","CXCL11","CXCL9","RIGI","DDX60","DHX58","EIF2AK2","EIF4E3","EPSTI1","FAS","FCGR1A","FGL2","FPR1","CMTR1","GBP4","GBP6","GCH1","GPR18","GZMA","HERC6","HIF1A","HLA-A","HLA-B","HLA-DMA","HLA-DQA1","HLA-DRB1","HLA-G","ICAM1","IDO1","IFI27","IFI30","IFI35","IFI44","IFI44L","IFIH1","IFIT1","IFIT2","IFIT3","IFITM2","IFITM3","IFNAR2","IL10RA","IL15","IL15RA","IL18BP","IL2RB","IL4R","IL6","IL7","IRF1","IRF2","IRF4","IRF5","IRF7","IRF8","IRF9","ISG15","ISG20","ISOC1","ITGB7","JAK2","KLRK1","LAP3","LATS2","LCP2","LGALS3BP","LY6E","LYSMD2","MARCHF1","METTL7B","MT2A","MTHFD2","MVP","MX1","MX2","MYD88","NAMPT","NCOA3","NFKB1","NFKBIA","NLRC5","NMI","NOD1","NUP93","OAS2","OAS3","OASL","OGFR","P2RY14","PARP12","PARP14","PDE4B","PELI1","PFKP","PIM1","PLA2G4A","PLSCR1","PML","PNP","PNPT1","HELZ2","PSMA2","PSMA3","PSMB10","PSMB2","PSMB8","PSMB9","PSME1","PSME2","PTGS2","PTPN1","PTPN2","PTPN6","RAPGEF6","RBCK1","RIPK1","RIPK2","RNF213","RNF31","RSAD2","RTP4","SAMD9L","SAMHD1","SECTM1","SELP","SERPING1","SLAMF7","SLC25A28","SOCS1","SOCS3","SOD2","SP110","SPPL2A","SRI","SSPN","ST3GAL5","ST8SIA4","STAT1","STAT2","STAT3","STAT4","TAP1","TAPBP","TDRD7","TNFAIP2","TNFAIP3","TNFAIP6","TNFSF10","TOR1B","TRAFD1","TRIM14","TRIM21","TRIM25","TRIM26","TXNIP","UBE2L6","UPP1","USP18","VAMP5","VAMP8","VCAM1","WARS1","XAF1","XCL1","ZBP1","ZNFX1")

merged4 = AddModuleScore(merged4, features = list(isg = isg))

merged4$Cluster1 %>% as_tibble() %>%
  mutate(donor = merged4$sample, time = merged4$time) %>% 
  group_by(donor, time) %>% summarise(ISG = mean(value))  %>% 
  inner_join(RNA_dat_summarised_bsm2) %>%
  filter(response != "Unclassified") %>%
  group_by(response, time) %>%
  mutate(c = cor(ISG, value)) %>%
  ggplot(aes(ISG, value)) +  facet_wrap(~response) + 
  geom_line(aes(group = time, colour = time), size = 2) +
  geom_line(aes(group = donor), linetype = "dashed") +
  geom_point(aes(fill= donor), shape = 21, size = 3) +
  scale_colour_manual(values = c("lightgrey", "grey", "darkgrey")) +
  scale_fill_manual(values = colour_donors) +
  theme_classic() +
  labs(x = "ISG signature expression", y = "RNA levels average.US.HIV.RNA.copies./10^6.18S") ->p;p
ggsave(p, filename = "../plots/Functional_data_ISG_RNA.pdf", width = 15, height = 9, units = "cm")

merged4$Cluster1 %>% as_tibble() %>%
  mutate(donor = merged4$sample, time = merged4$time) %>% 
  group_by(donor, time) %>% summarise(ISG = mean(value))  %>% 
  inner_join(RNA_dat_summarised_bsm2) %>% 
  filter(response != "Unclassified") %>%
  group_by(response) %>%
  summarise(c = cor(ISG, value)) 
```
Correlation of ISG with RNA for single DNA for multiple
in responders -> RNA up at single doese and correlated with ISG
Over time the DNA then goes down

split timepoints into +1 +8

```{r}
DNA_dat %>% select(-"Average.number.HIV.DNA.copies", -time) %>% 
  mutate(Timepoint = recode(Timepoint,
                            "C1 D+1" = "C1D+1",
                            "C1 D8" = "C1D8",
                            "C4 D-1" = "C4D11", 
                            "C13 D-1" = "C13D-1")) %>% 
  rbind(DNA_dat4  %>% select(-time) %>% dplyr::rename("average.HIV.copies/10^6.CCR5" = "average.HIV.copies/10^.CCR5")) %>% 
  mutate(donor = factor(donor, levels = c("Donor 38", "Donor 49", "Donor 56",
                                            "Donor 41",  "Donor 62", "Donor 64", "Donor 48", "Donor 54"))) %>% 
  group_by(donor, Timepoint, response) %>%
  summarise(value = mean(`average.HIV.copies/10^6.CCR5`)) -> DNA_dat_summarised4


DNA_dat %>% select(-"Average.number.HIV.DNA.copies", -Timepoint) %>% 
  rbind(DNA_dat4  %>% select(-Timepoint) %>% dplyr::rename("average.HIV.copies/10^6.CCR5" = "average.HIV.copies/10^.CCR5")) %>% 
  mutate(donor = factor(donor, levels = c("Donor 38", "Donor 49", "Donor 56",
                                            "Donor 41",  "Donor 62", "Donor 64", "Donor 48", "Donor 54"))) %>% 
  group_by(donor, time, response) %>%
  summarise(value = mean(`average.HIV.copies/10^6.CCR5`)) -> DNA_dat_summarised4_bsm

merged4$Cluster1 %>% as_tibble() %>%
  mutate(donor = merged4$sample, Timepoint = merged4$Timepoint) %>% 
  mutate(Timepoint = recode(Timepoint,
                       "Baseline" = "BL",  "C01D-1" = "BL",
                       "C01D+1" = "C1D+1", "C01D+8" = "C1D8",
                       "C04D-1" = "C4D-1", "C04D+1" = "C4D+1")) %>%
  group_by(donor, Timepoint) %>% summarise(ISG = mean(value))  %>%
  inner_join(DNA_dat_summarised4) %>% 
  filter(response != "Unclassified") %>%
  ggplot(aes(ISG, value)) + geom_point(aes(color= Timepoint)) + facet_wrap(~response) + geom_line(aes(group = Timepoint, colour = Timepoint)) +
  geom_line(aes(group = donor), linetype = "dashed")

merged4$Cluster1 %>% as_tibble() %>%
  mutate(donor = merged4$sample, Timepoint = merged4$Timepoint) %>% 
  mutate(Timepoint = recode(Timepoint,
                       "Baseline" = "BL",  "C01D-1" = "BL",
                       "C01D+1" = "C1D+1", "C01D+8" = "C1D8",
                       "C04D-1" = "C4D-1", "C04D+1" = "C4D+1")) %>%
  group_by(donor, Timepoint) %>% summarise(ISG = mean(value))  %>%
  inner_join(DNA_dat_summarised4) %>% 
  filter(response != "Unclassified") %>%
  ggplot(aes(ISG, value)) + geom_point() + facet_wrap(~response) + 
  geom_smooth()

merged4$Cluster1 %>% as_tibble() %>%
  mutate(donor = merged4$sample, Timepoint = merged4$Timepoint) %>% 
  mutate(Timepoint = recode(Timepoint,
                       "Baseline" = "BL",  "C01D-1" = "BL",
                       "C01D+1" = "C1D+1", "C01D+8" = "C1D8",
                       "C04D-1" = "C4D-1", "C04D+1" = "C4D+1")) %>%
  group_by(donor, Timepoint) %>% summarise(ISG = mean(value))  %>%
  inner_join(DNA_dat_summarised4) %>% 
  filter(response != "Unclassified") %>% 
  group_by(response) %>%
  summarise(cor = cor(ISG, value))
```
```{r}
merged4$Cluster1 %>% as_tibble() %>%
  mutate(donor = merged4$sample, time = merged4$time) %>% 
  group_by(donor, time) %>% summarise(ISG = mean(value))  %>% 
  inner_join(DNA_dat_summarised4_bsm) %>%
  filter(response != "Unclassified") %>%
  group_by(response, time) %>%
  mutate(c = cor(ISG, value)) %>%
  ggplot(aes(ISG, value)) + 
  geom_line(aes(group = time, colour = time), size = 2) +
  geom_point(aes(fill= donor), shape = 21, size = 3) + facet_wrap(~response) + 
  geom_line(aes(group = donor), linetype = "dashed") +
  scale_colour_manual(values = c("lightgrey", "grey", "darkgrey")) +
  scale_fill_manual(values = colour_donors) +
  theme_classic() +
  labs(x = "ISG signature expression", y = "DNA levels average.HIV.copies/10^6.CCR5")->p;p
ggsave(p, filename = "../plots/Functional_data_ISG_DNA.pdf", width = 15, height = 9, units = "cm")
```

```{r}
merged4@assays$SCT@data["MX1",] %>% as_tibble() %>%
  mutate(donor = merged4$sample, time = merged4$time) %>% 
  group_by(donor, time) %>% summarise(ISG = mean(value))  %>% 
  inner_join(DNA_dat_summarised4_bsm) %>%
  filter(response != "Unclassified") %>%
  group_by(response, time) %>%
  mutate(c = cor(ISG, value)) %>%
  ggplot(aes(ISG, value)) + 
  geom_line(aes(group = time, colour = time), size = 2) +
  geom_point(aes(fill= donor), shape = 21, size = 3) + facet_wrap(~response) + 
  geom_line(aes(group = donor), linetype = "dashed") +
  scale_colour_manual(values = c("lightgrey", "grey", "darkgrey")) +
  scale_fill_manual(values = colour_donors) +
  theme_classic() +
  labs(x = "MX1 expression", y = "DNA levels average.HIV.copies/10^6.CCR5") ->p;p
ggsave(p, filename = "../plots/Functional_data_MX1_DNA.pdf", width = 15, height = 9, units = "cm")
```

```{r}
merged4@assays$SCT@data["STAT1",] %>% as_tibble() %>%
  mutate(donor = merged4$sample, time = merged4$time) %>% 
  group_by(donor, time) %>% summarise(ISG = mean(value))  %>% 
  inner_join(DNA_dat_summarised4_bsm) %>%
  filter(response != "Unclassified") %>%
  group_by(response, time) %>%
  mutate(c = cor(ISG, value)) %>%
  ggplot(aes(ISG, value)) + 
  geom_line(aes(group = time, colour = time), size = 2) +
  geom_point(aes(fill= donor), shape = 21, size = 3) + facet_wrap(~response) + 
  geom_line(aes(group = donor), linetype = "dashed") +
  scale_colour_manual(values = c("lightgrey", "grey", "darkgrey")) +
  scale_fill_manual(values = colour_donors) +
  theme_classic() +
  labs(x = "STAT1 expression", y = "DNA levels average.HIV.copies/10^6.CCR5")->p;p
ggsave(p, filename = "../plots/Functional_data_STAT1_DNA.pdf", width = 15, height = 9, units = "cm")
```


## Correlate Ratios with genes

```{r}
ggplot(ratio_dat, aes(time, ratio)) +
  geom_point() + facet_wrap(~donor, scales = "free_y")
```

```{r}
plot_gene_correlation = function(g) {
  merged4@meta.data %>% as_tibble() %>%
    mutate(exp = merged4@assays$SCT@data[g,]) %>%
    group_by(sample, time) %>%
    summarise(exp = mean(exp)) %>%
    inner_join(ratio_dat , by = c("sample" = "donor", "time" ="time")) %>%
    ggplot(aes(ratio, exp)) + geom_point(aes(colour = time, shape = sample)) + geom_line(aes(group = time, colour = time))
}
```


```{r}
cors = data.frame()
for(g in row.names(merged4@assays$SCT@data)){
  merged4@meta.data %>% as_tibble() %>%
    mutate(exp = merged4@assays$SCT@data[g,]) %>%
    filter(time == "No dose") %>%
    group_by(sample) %>%
    summarise(exp = mean(exp)) %>%
    inner_join(ratio_dat %>% filter(time == "No dose"), by = c("sample" = "donor")) ->cd
  cors = rbind(cors, c(g, cor(cd$ratio, cd$exp), mean(cd$exp)))
  
}
names(cors) = c("gene", "cor", "mean_exp")
cors_c1 %>% as_tibble() %>% arrange(desc(cor)) %>% head(n=15)

inner_join(cors, cors_c1, by = "gene") %>% as_tibble %>% 
  filter(!is.na(cor.x), !is.na(cor.y)) %>% mutate(diff = as.numeric(cor.x)-as.numeric(cor.y)) %>% 
  filter(mean_exp.x > 1 | mean_exp.y > 1) %>%
  # filter(abs(as.numeric(cor.x)) < 0.1) %>%
  arrange(-abs(diff))
```
 
## Ratios over time
 
```{r}
ggplot(RNA_dat_summarised2, aes(time, value)) +
  geom_point()+
  facet_wrap(~donor)
```
### Correlate baseline RNA with GEX

```{r}
plot_gene_correlation2 = function(g) {
  merged4@meta.data %>% as_tibble() %>%
    mutate(exp = merged4@assays$SCT@data[g,]) %>%
    group_by(sample, time) %>%
    summarise(exp = mean(exp)) %>%
    inner_join(RNA_dat_summarised2 , by = c("sample" = "donor", "time" ="time")) %>%
    ggplot(aes(value, exp)) + geom_point(aes(colour = time, shape = sample)) + geom_line(aes(group = time, colour = time))
}
```
```{r}
cors = data.frame()
for(g in row.names(merged4@assays$SCT@data)){
  merged4@meta.data %>% as_tibble() %>%
    mutate(exp = merged4@assays$SCT@data[g,]) %>%
    filter(time == "No dose") %>%
    group_by(sample) %>%
    summarise(exp = mean(exp)) %>%
    inner_join(RNA_dat_summarised2 %>% filter(time == "No dose"), by = c("sample" = "donor")) ->cd
  cors = rbind(cors, c(g, cor(cd$value, cd$exp), mean(cd$exp)))
  
}
names(cors) = c("gene", "cor", "mean_exp")
cors %>% as_tibble() %>% arrange(desc(cor)) %>% head(n=15)

write_csv(x = cors %>% as_tibble() %>% arrange(desc(cor)), file = "../output/correlations_RNA_baseline_GEX.csv")
```

## Correlate ratio between BL and C1

```{r}
plot_gene_correlation3 = function(g) {
  gene_ratios %>% inner_join(rna_ratios, by = c("sample" = "donor")) %>%
    filter(gene == g) %>%
    ggplot(aes(ratio.x, ratio.y))+ geom_point() + 
    geom_text_repel(aes(label = sample))
}
```
```{r}
merged4$sample_time = paste(merged4$sample, merged4$time)
PB = AverageExpression(merged4, group.by = "sample_time", features = row.names(merged4@assays$SCT@data), assays = "SCT")
PB$SCT %>% as_tibble(rownames = "gene") %>% 
  pivot_longer(-gene) %>%
  filter(!grepl("Multi", name)) %>%
  mutate(sample = paste("Donor" , limma::strsplit2(name, split = " ")[,2])) %>%
  group_by(gene, sample) %>% summarise(ratio = value[2]/value[1]) -> gene_ratios

RNA_dat_summarised2 %>% filter(time != "Multi dose") %>%
  pivot_wider(names_from = "time", values_from = value) %>%
  mutate(ratio = `Single dose`/`No dose`) %>% filter(!is.na(ratio)) -> rna_ratios

gene_ratios %>% inner_join(rna_ratios, by = c("sample" = "donor")) %>%
  group_by(gene) %>%
  summarise(c = cor(ratio.x, ratio.y)) -> cors2

cors2 %>% arrange(desc(c))

# cors2 = data.frame()
# for(g in row.names(merged4@assays$SCT@data)){
#   PB$SCT[g,] %>%
#     group_by(sample, time) %>%
#     summarise(exp = mean(exp)) %>%
#     pivot_wider(names_from = "time", values_from = "exp") %>%
#     mutate(R = `Single dose`/`No dose`) %>%
#     inner_join(rna_ratios, by = c("sample" = "donor")) ->cd
#   cors2 = rbind(cors2, c(g, cor(cd$ratio, cd$R), mean(cd$exp)))
#   
# }
```

## RNA Change correlation
```{r}
plot_gene_correlation4 = function(g) {
  PB$SCT %>% as_tibble(rownames = "gene") %>% 
    filter(gene == g) %>%
    pivot_longer(-gene) %>%
    
    filter(grepl("No dose", name)) %>% 
    mutate(sample = paste("Donor" , limma::strsplit2(name, split = " ")[,2])) %>%
    inner_join(rna_ratios, by = c("sample" = "donor")) %>%
    ggplot(aes(ratio, value))+ geom_point() + 
    geom_text_repel(aes(label = sample))
}
```

```{r}
plot_gene_correlation_rna = function(g) {
  PB$SCT %>% as_tibble(rownames = "gene") %>% 
    filter(gene == g) %>%
    pivot_longer(-gene) %>% 
    mutate(sample = paste("Donor" , limma::strsplit2(name, split = " ")[,2])) %>%
    mutate(time = limma::strsplit2(name, split = " ")[,3]) %>%
    inner_join(rna_ratios, by = c("sample" = "donor")) %>%
    ggplot(aes(ratio, value, color = time))+ geom_point() +
    geom_line(aes(group = time)) +
    geom_text_repel(aes(label = sample)) +
    ggtitle(g)
}
```

```{r}
PB$SCT %>% as_tibble(rownames = "gene") %>% 
  pivot_longer(-gene) %>%
  filter(grepl("No dose", name)) %>% 
  mutate(sample = paste("Donor" , limma::strsplit2(name, split = " ")[,2])) %>%
  inner_join(rna_ratios, by = c("sample" = "donor")) %>% 
  group_by(gene) %>% summarise(cor = cor(value, ratio), mean = mean(value)) %>%
  arrange(desc(cor)) -> cors3
```

```{r}
cors3 %>% filter(cor>0.8 | cor< -0.8, mean>0.5) %>% 
  left_join(PB$SCT %>% as_tibble(rownames = "gene")) %>%
  pivot_longer(cols = starts_with("Donor")| starts_with("PBMC")) %>%
  filter(grepl("Donor", name), grepl("No dose", name)) %>%
  group_by(gene) %>% arrange(gene, value) %>% mutate(r = ifelse(cor>0,1:6, 6:1)) %>% 
  ungroup() %>%
  count(name, r) %>%
  mutate(name = factor(name)) %>% mutate(name = factor(name, levels = levels(name)[c(6, 2, 4, 3, 5, 1)])) %>%
  ggplot(aes(r, name)) +
  geom_tile(aes(fill = n)) +
  geom_text(aes(label = n), colour = "white") 
```

## DNA Ratios Prediction
```{r}
plot_gene_correlation5 = function(g) {
  PB$SCT %>% as_tibble(rownames = "gene") %>% 
    filter(gene == g) %>%
    pivot_longer(-gene) %>%
    
    filter(grepl("No dose", name)) %>% 
    mutate(sample = paste("Donor" , limma::strsplit2(name, split = " ")[,2])) %>%
    inner_join(dna_ratios, by = c("sample" = "donor")) %>%
    ggplot(aes(ratio, value))+ geom_point() + 
    geom_text_repel(aes(label = sample))
}
```

```{r}
DNA_dat_summarised4 %>% filter(time != "Multi dose") %>%
  pivot_wider(names_from = "time", values_from = value) %>%
  mutate(ratio = `Single dose`/`No dose`) %>% filter(!is.na(ratio)) -> dna_ratios


PB$SCT %>% as_tibble(rownames = "gene") %>% 
  pivot_longer(-gene) %>%
  filter(grepl("No dose", name)) %>% 
  mutate(sample = paste("Donor" , limma::strsplit2(name, split = " ")[,2])) %>%
  inner_join(dna_ratios, by = c("sample" = "donor")) %>% 
  group_by(gene) %>% summarise(cor = cor(value, ratio), mean = mean(value)) %>%
  arrange(desc(cor)) -> cors4

head(cors4 %>% arrange((cor)))
```
```{r}
cors4 %>% filter(cor>0.8 | cor< -0.8, mean>0.5) %>% 
  left_join(PB$SCT %>% as_tibble(rownames = "gene")) %>%
  pivot_longer(cols = starts_with("Donor")| starts_with("PBMC")) %>%
  filter(grepl("Donor", name), grepl("No dose", name)) %>%
  group_by(gene) %>% arrange(gene, value) %>% mutate(r = 1:6) %>% 
  mutate(cordir = ifelse(cor>0, "pos", "neg")) %>%
  ungroup() %>%
  count(name, r, cordir) %>%
  ggplot(aes(r, name)) +
  geom_tile(aes(fill = n)) +
  geom_text(aes(label = n), colour = "white") +
  facet_wrap(~cordir)

cors4 %>% filter(cor>0.8 | cor< -0.8, mean>0.5) %>% 
  left_join(PB$SCT %>% as_tibble(rownames = "gene")) %>%
  pivot_longer(cols = starts_with("Donor")| starts_with("PBMC")) %>%
  filter(grepl("Donor", name), grepl("No dose", name)) %>%
  group_by(gene) %>% arrange(gene, value) %>% mutate(r = ifelse(cor>0,1:6, 6:1)) %>% 
  ungroup() %>%
  count(name, r) %>%
  mutate(name = factor(name)) %>% mutate(name = factor(name, levels = levels(name)[c(1,2,6,3,5,4)])) %>%
  ggplot(aes(r, name)) +
  geom_tile(aes(fill = n)) +
  geom_text(aes(label = n), colour = "white") 
```

## RNA/DNA Ratios Prediction


```{r}
RNA_dat_summarised2 %>% inner_join(DNA_dat_summarised4, by = c("donor", "time", "response")) %>% 
  mutate(ratio = value.x/value.y) %>% 
  filter(!is.infinite(ratio)) %>%
  select(-value.x, -value.y) %>%
  pivot_wider(names_from = "time", values_from = "ratio") %>%
  mutate(ratio = `Single dose`/`No dose`) %>% 
  filter(!is.na(ratio)) ->ratio_dat



PB$SCT %>% as_tibble(rownames = "gene") %>% 
  pivot_longer(-gene) %>%
  filter(grepl("No dose", name)) %>% 
  mutate(sample = paste("Donor" , limma::strsplit2(name, split = " ")[,2])) %>%
  inner_join(ratio_dat, by = c("sample" = "donor")) %>% 
  group_by(gene) %>% summarise(cor = cor(value, ratio), mean = mean(value)) %>%
  arrange(desc(cor)) -> cors5

head(cors5 %>% arrange((cor)))
head(cors5 %>% arrange(desc(mean),desc(cor))) 
```
```{r}
cors5 %>% filter(cor>0.8 | cor< -0.8, mean>0.5) %>%  
  left_join(PB$SCT %>% as_tibble(rownames = "gene")) %>%
  pivot_longer(cols = starts_with("Donor")| starts_with("PBMC")) %>%
  filter(grepl("Donor", name), grepl("No dose", name)) %>%
  group_by(gene) %>% arrange(gene, value) %>% mutate(r = 1:6) %>% 
  mutate(cordir = ifelse(cor>0, "pos", "neg")) %>%
  ungroup() %>%
  count(name, r, cordir) %>%
  ggplot(aes(r, name)) +
  geom_tile(aes(fill = n)) +
  geom_text(aes(label = n), colour = "white") +
  facet_wrap(~cordir)

cors5 %>% filter(cor>0.8 | cor< -0.8, mean>0.5) %>% 
  left_join(PB$SCT %>% as_tibble(rownames = "gene")) %>%
  pivot_longer(cols = starts_with("Donor")| starts_with("PBMC")) %>%
  filter(grepl("Donor", name), grepl("No dose", name)) %>%
  group_by(gene) %>% arrange(gene, value) %>% mutate(r = ifelse(cor>0,1:6, 6:1)) %>% 
  ungroup() %>%
  count(name, r) %>%
  mutate(name = factor(name)) %>% mutate(name = factor(name, levels = levels(name)[c(6,4,5,2,3,1)])) %>%
  ggplot(aes(r, name)) +
  geom_tile(aes(fill = n)) +
  geom_text(aes(label = n), colour = "white") 
```
## Pseudo Bulk
```{r}
library(edgeR)
pseudo = AggregateExpression(merged4, group.by = "sample_time", assays = "RNA")

dge = DGEList(pseudo$RNA)
dge = dge[,grepl("No dose", row.names(dge$samples))]

summary(filterByExpr(dge, min.count = 3))
dge = dge[filterByExpr(dge, min.count = 3),]

dge = calcNormFactors(dge)

plotMDS(dge)

dge$samples$order = c(6,4,5,2,3,1)
```

```{r}
dge2 = dge[,-5]
design = model.matrix(~1+ratio_dat$ratio[c(1, 4, 2, 3, 5)]+
                        dna_ratios$ratio[c(1,4,2,3,5)] +
                        rna_ratios$ratio[c(1,4,2,3,5)])
colnames(design) = c("Intercept", "ratio", "dna_ratio", "rna_ratio")

dge2 = estimateDisp(dge2, design, robust = T)

fit = glmFit(dge2, design)
lrt_rna_dna_ratio = glmLRT(fit, coef = 2)
lrt_dna_ratio = glmLRT(fit, coef = 3)
lrt_rna_ratio = glmLRT(fit, coef = 4)

summary(decideTests.DGELRT(lrt_rna_dna_ratio, p.value = 0.2))
summary(decideTests.DGELRT(lrt_dna_ratio, p.value = 0.2))
summary(decideTests.DGELRT(lrt_rna_ratio, p.value = 0.2))

topTags(lrt_rna_ratio, n=50)

data.frame(exp = cpm(dge2,log = T)["SELL",], ratio = rna_ratios$ratio[c(1, 4, 2, 3, 5)]) %>%
  ggplot(aes(ratio, exp)) + geom_point()
```
### KEGG
```{r}
library(org.Hs.eg.db)
entrez = mapIds(org.Hs.eg.db, row.names(dge2), "SYMBOL", column = "ENTREZID")

kegg_rna = kegga.DGELRT(lrt_rna_ratio, species = "Hs", geneid = entrez, FDR = 0.2)
topKEGG(kegg_rna)

kegg_dna = kegga.DGELRT(lrt_dna_ratio, species = "Hs", geneid = entrez, FDR = 0.2)
topKEGG(kegg_dna)
```

### Hallmark

```{r}
mh = read.table("~/projects/references/human/Hallmark_Gene_Sets/h.all.v2023.2.Hs.symbols.gmt", sep = "%")

data = list()

con = file("~/projects/references/human/Hallmark_Gene_Sets/h.all.v2023.2.Hs.symbols.gmt", "r")
  while ( TRUE ) {
    line = readLines(con, n = 1)
    if ( length(line) == 0 ) {
      break
    }
    split = strsplit2(line, split = "\t")
    #print(split)
    data[[split[1]]] = split[-c(1:2)]
  }

  close(con)
  
print("blank")
```

```{r}

roasts_rna_ratio = lapply(data, function(x) roast(dge2, design = design, index = x,contrast = 4 ))

rbind(roasts_rna_ratio[[2]]$p.value[,2], roasts_rna_ratio[[1]]$p.value[,2])

p_vals = lapply(roasts_rna_ratio, function(x) x$p.value[,2])
p_vals_df = purrr::reduce(p_vals, rbind)

row.names(p_vals_df) = names(p_vals)
colnames(p_vals_df) = c("Down", "Up", "UpOrDown", "Mixed")

p_vals_df %>% as_tibble(rownames = "pathway") %>% arrange(Down)
p_vals_df %>% as_tibble(rownames = "pathway") %>% arrange(Up)
```
```{r}
roasts_dna_ratio = lapply(data, function(x) roast(dge2, design = design, index = x, contrast = 3))


p_vals_dna = lapply(roasts_dna_ratio, function(x) x$p.value[,2])
p_vals_df_dna = purrr::reduce(p_vals_dna, rbind)

row.names(p_vals_df_dna) = names(p_vals_dn)
colnames(p_vals_df_dna) = c("Down", "Up", "UpOrDown", "Mixed")

p_vals_df_dna %>% as_tibble(rownames = "pathway") %>% arrange(Down)
p_vals_df_dna %>% as_tibble(rownames = "pathway") %>% arrange(Up)
```
```{r}

roasts_rna_dna_ratio = lapply(data, function(x) roast(dge2, design = design, index = x,contrast = 2 ))

rbind(roasts_rna_dna_ratio[[2]]$p.value[,2], roasts_rna_dna_ratio[[1]]$p.value[,2])

p_vals = lapply(roasts_rna_dna_ratio, function(x) x$p.value[,2])
p_vals_df_rna_dna = purrr::reduce(p_vals, rbind)

row.names(p_vals_df_rna_dna) = names(p_vals)
colnames(p_vals_df_rna_dna) = c("Down", "Up", "UpOrDown", "Mixed")

p_vals_df_rna_dna %>% as_tibble(rownames = "pathway") %>% arrange(Down)
p_vals_df_rna_dna %>% as_tibble(rownames = "pathway") %>% arrange(Up)
```

```{r}
p_vals_df_rna_dna %>% as_tibble(rownames = "pathway") %>%
  mutate(Up_RNA = p_vals_df[,2], Dn_RNA = p_vals_df[,1],
        Up_DNA = p_vals_df_dna[,2], Dn_DNA = p_vals_df_dna[,1] ) %>%
  filter( (Up<0.1 | Down <0.1) & (Up_RNA<0.1 | Dn_RNA<0.1)  & (Up_DNA<0.1 | Dn_DNA < 0.1)) %>%
  dplyr::select (-UpOrDown, - Mixed)
```

### PB DGE Per Cluster
```{r}
merged4$gg = paste(merged4$sample, merged4$time, merged4$seurat_clusters)
pseudo = AggregateExpression(merged4, group.by = "gg", assays = "RNA")

dge = DGEList(pseudo$RNA)
dge = dge[,grepl("No dose", row.names(dge$samples))]

summary(filterByExpr(dge, min.count = 1, min.total.count = 1, min.prop = 0))
dge = dge[filterByExpr(dge, min.count = 1, min.total.count = 1, min.prop = 0),]

dge = calcNormFactors(dge)

plotMDS(dge)

dge$samples$order = c(6,4,5,2,3,1)
```
```{r}
dge2 = dge[,grepl(" 2$", row.names(dge$samples))]
design = model.matrix(~1+rna_ratios$ratio[c(1,4,2,3,5)])
colnames(design) = c("Intercept", "rna_ratio")

dge2 = estimateDisp(dge2, design, robust = T)

fit = glmFit(dge2, design)
lrt_rna_ratio = glmLRT(fit, coef = 2)


summary(decideTests.DGELRT(lrt_rna_ratio, p.value = 0.2))

topTags(lrt_rna_ratio, n=50)

data.frame(exp = cpm(dge2,log = T)["SELL",], ratio = rna_ratios$ratio[c(1, 4, 2, 3, 5)]) %>%
  ggplot(aes(ratio, exp)) + geom_point()
```




```{r}
merged4 = AddModuleScore(merged4, features = list(kras = data$HALLMARK_KRAS_SIGNALING_DN), name = "HALLMARK_KRAS_SIGNALING_D")

plotto_signature_scoring_plot(merged4, sig = "HALLMARK_KRAS_SIGNALING_D1", reduction = "harmony_umap", size = 1)

VlnPlot(merged4[,!grepl("PBMC", merged4$sample) &
                  !grepl("Multi", merged4$time)], "HALLMARK_KRAS_SIGNALING_D1", 
        group.by = "sample_time")

VlnPlot(merged4[,merged4$time == "No dose" ], "HALLMARK_KRAS_SIGNALING_D1", 
        group.by = "seurat_clusters")

merged4@meta.data %>% as_tibble() %>%
  filter(time == "No dose") %>%
  group_by(seurat_clusters) %>% mutate(m=median(HALLMARK_KRAS_SIGNALING_D1)) -> pdat;

  ggplot(pdat,aes(reorder(seurat_clusters, m), HALLMARK_KRAS_SIGNALING_D1)) +
  geom_violin(draw_quantiles = 0.5)

ggplot(pdat %>% filter(seurat_clusters == 6), aes(sample, HALLMARK_KRAS_SIGNALING_D1)) +
  geom_violin()

pdat %>% count(sample, seurat_clusters) %>%
  group_by(sample) %>% mutate(f = n/sum(n)*100) %>%
  ggplot(aes(sample, f)) + geom_point() + facet_wrap(~seurat_clusters)

merged4@meta.data %>% as_tibble() %>%
    filter(time == "No dose") %>%
  group_by(sample) %>% mutate(m=mean(HALLMARK_KRAS_SIGNALING_D1)) %>%
  ggplot(aes(reorder(sample, m), HALLMARK_KRAS_SIGNALING_D1)) +
  geom_violin(draw_quantiles = 0.5)

merged4@meta.data %>% as_tibble() %>%
    filter(time == "No dose") %>%
  group_by(sample) %>% mutate(tot=n()) %>%
  group_by(seurat_clusters, sample) %>% summarise(m=mean(HALLMARK_KRAS_SIGNALING_D1), f = n()/tot[1]*100) %>%
  filter(seurat_clusters == 6) %>%
  ggplot(aes(m,f)) +
  geom_point(aes(color=seurat_clusters))
```
```{r}
merged4 = SetIdent(merged4, value = "seurat_clusters")
merged4 = FindSubCluster(merged4, cluster = 1, subcluster.name = "C1_sub" , graph.name = "SCT_nn")

VlnPlot(merged4[,!grepl("PBMC", merged4$sample) &
                  grepl("No dose", merged4$time) &
                  merged4$seurat_clusters == 6], 
        "HALLMARK_KRAS_SIGNALING_D1", 
        group.by = "sample")

VlnPlot(merged4[,!grepl("PBMC", merged4$sample) &
                  grepl("No dose", merged4$time) &
                  merged4$C1_sub == "1_2"], 
        "HALLMARK_KRAS_SIGNALING_D1", 
        group.by = "sample")
```
## Cluster Proportions vs virus



```{r}
#merged4 = FindNeighbors(merged4, reduction = "harmony", dims = 1:30)
#merged4 = FindClusters(merged4, resolution = 1)

merged4@meta.data %>% as_tibble() %>%
  count(sample, time, seurat_clusters) %>%
  group_by(sample, time) %>%
  mutate(f = n/sum(n)*100) %>%
  filter(!grepl("PBMC", sample)) -> frequencies

RNA_dat_summarised2 %>% inner_join(DNA_dat_summarised4, by = c("donor", "time", "response")) %>% 
  mutate(ratio = value.x/value.y) %>% 
  filter(!is.infinite(ratio)) %>%
  select(-value.x, -value.y) %>%
  pivot_wider(names_from = "time", values_from = "ratio") %>%
  mutate(ratio = `Single dose`/`No dose`) %>% 
  filter(!is.na(ratio)) ->rna_over__dnaratios

frequencies %>%
  filter(time == "No dose") %>%
  inner_join(rna_over__dnaratios, by = c("sample" = "donor")) %>%
  group_by(seurat_clusters) %>% 
  summarise(cor = cor(ratio, f))

frequencies %>%
  filter(time == "No dose") %>%
  inner_join(rna_over__dnaratios, by = c("sample" = "donor")) %>%
  ggplot(aes(f, ratio)) + geom_point() + facet_wrap(~seurat_clusters) 
```
```{r}
DNA_dat_summarised4 %>% filter(time != "Multi dose") %>%
  pivot_wider(names_from = "time", values_from = value) %>%
  mutate(ratio = `Single dose`/`No dose`) %>% filter(!is.na(ratio)) -> dna_ratios

frequencies %>%
  filter(time == "No dose") %>%
  inner_join(dna_ratios, by = c("sample" = "donor")) %>%
  group_by(seurat_clusters) %>% 
  summarise(cor = cor(ratio, f))

frequencies %>%
  filter(time == "No dose") %>%
  inner_join(dna_ratios, by = c("sample" = "donor")) %>%
  ggplot(aes(f, ratio)) + geom_point() + facet_wrap(~seurat_clusters)
```
### Load Updated Metadata
```{r}
library(openxlsx)
RNA_dat3 = readWorkbook("../data/Gubser_aPD-1 Suppl material DEC25.xlsx", sheet = 7, startRow = 2, colNames = T)
RNA_dat3 = RNA_dat3[!is.na(RNA_dat3$`average.US.HIV.copies/10^6.18S`),]
RNA_dat3 %>% filter(grepl("C1D", Timepoint) | grepl("Baseline", Timepoint)) %>%
  select(Study.ID, Timepoint, "average.US.HIV.copies/10^6.18S") %>% 
  fill(Study.ID, .direction = "down") %>%
  mutate(Timepoint = recode(Timepoint, "C1D-1=Baseline" = "Baseline", "C1D-1" = "Baseline",
                            "C1D-1 =Baseline" = "Baseline")) %>%
  mutate(Timepoint  = ifelse(Timepoint == "Baseline", "Baseline", "Single")) %>%
  group_by(Timepoint, Study.ID) %>% summarise("average.US.HIV.copies/10^6.18S" = mean(`average.US.HIV.copies/10^6.18S`)) -> RNA_dat_summarised3

RNA_dat_summarised2 %>%
  mutate(Study.ID = donor_mapping[as.character(donor)]) %>% 
  mutate(Timepoint = recode(time, "No dose" = "Baseline", "Single dose" = "Single")) %>%
  left_join(RNA_dat_summarised3) %>%
  mutate(value = `average.US.HIV.copies/10^6.18S`) %>% select(donor, time, value) -> RNA_dat_summarised_updated
```

```{r}
library(openxlsx)
# DNA_data3 = readWorkbook("../data/Gubser_aPD-1 Suppl material DEC25.xlsx", sheet = 7, startRow = 2, colNames = T)
# DNA_data3 = DNA_data3[!is.na(DNA_data3$`average.US.HIV.copies/10^6.18S`),]
# DNA_data3 %>% filter(grepl("C1D", Timepoint) | grepl("Baseline", Timepoint)) %>%
#   select(Study.ID, Timepoint, `average.US.HIV.copies/10^6.18S`) %>%
#   fill(Study.ID, .direction = "down") %>%
#   mutate(Timepoint = recode(Timepoint, "C1D-1=Baseline" = "Baseline", "C1D-1" = "Baseline")) %>%
#   mutate(Timepoint  = ifelse(Timepoint == "Baseline", "Baseline", "Single")) %>%
#   group_by(Timepoint, Study.ID) %>% summarise(`average.US.HIV.copies/10^6.18S` = mean(`average.US.HIV.copies/10^6.18S`)) -> RNA_dat_summarised3


  
```

```{r}
library(ggrepel)
# RNA_dat_summarised3 %>%
#   pivot_wider(names_from = Timepoint, values_from = "average.US.HIV.copies/10^6.18S") %>% 
#   mutate(ratio = Single/Baseline) %>% filter(!is.na(ratio)) -> rna_ratios
RNA_dat_summarised2 %>%
  pivot_wider(names_from = time, values_from = "value") %>% 
  mutate(ratio = `Single dose`/`No dose`) %>% filter(!is.na(ratio)) -> rna_ratios

RNA_dat_summarised_updated %>%
  pivot_wider(names_from = time, values_from = value) %>% 
  mutate(ratio = `Single dose`/`No dose`) %>% filter(!is.na(ratio)) -> rna_ratios

DNA_dat_summarised3 %>%
  pivot_wider(names_from = Timepoint, values_from = "average.US.HIV.copies/10^6.18S") %>% 
  mutate(ratio = Single/Baseline) %>% filter(!is.na(ratio)) -> dna_ratios



frequencies %>%
  filter(time == "No dose") %>%
  mutate(Study.ID = donor_mapping[as.character(sample)]) %>%
  inner_join(rna_ratios) %>%
  group_by(seurat_clusters) %>% 
  summarise(cor = cor(ratio, f))

frequencies %>%
  filter(time == "No dose") %>%
  inner_join(rna_ratios, by = c("sample" = "donor"))  -> fdat

lms = lapply(as.list(unique(fdat$seurat_clusters)), function(x) 
  lm(fdat$ratio[fdat$seurat_clusters == x] ~ fdat$f[fdat$seurat_clusters == x]))

stats = lapply(lms, function(x) summary(x)$coefficients)

pvals = purrr::reduce(stats, rbind) %>% as_tibble(rownames = "s") %>%
  filter(s != "(Intercept)")
intercepts = purrr::reduce(stats, rbind) %>% as_tibble(rownames = "s") %>%
  filter(s == "(Intercept)")



fdat %>%
  mutate(direction = ratio > 1) %>%
  mutate(seurat_clusters = as.numeric(as.character(seurat_clusters))) %>%
  left_join(data.frame(seurat_clusters = 0:8, int = intercepts$Estimate, slope = pvals$Estimate)) %>%
  group_by(seurat_clusters, direction) %>%
  summarise(value = case_when(direction == TRUE & slope[1] > 0 ~ min(f),
                              direction == TRUE & slope[1] < 0 ~ max(f),
                              direction == FALSE & slope[1] > 0 ~ max(f),
                              .default = min(f)),
            slope = ifelse(slope[1] > 0, "pos", "neg")) %>%
  distinct() %>%
  ungroup() %>% mutate( direction = ifelse(direction, "up", "down")) %>%
  pivot_wider(names_from = direction, values_from = value) %>% 
  filter(slope == "pos" & up>down | slope =="neg" & up<down) %>%
  mutate(intercept = (up+down)/2) -> divdata
  

frequencies %>%
  filter(time == "No dose") %>%
  inner_join(rna_ratios, by = c("sample" = "donor"))
  
ggplot(fdat, aes(f, ratio)) + geom_point() + facet_wrap(~seurat_clusters) + 
  geom_abline(data = data.frame(seurat_clusters = 0:8, int = intercepts$Estimate, slope = pvals$Estimate),
              aes(intercept = int, slope = slope), colour = "darkblue", linetype = "dashed") +
    geom_label(data = data.frame(seurat_clusters = 0:8, l = paste("p =", round(pvals$`Pr(>|t|)`, 3))), 
              aes(label = l), x = 37, y = 1.1, colour = "darkblue") +
    theme_classic() +
    geom_hline(yintercept = 1) +
    geom_vline(data = divdata, aes(xintercept = intercept), colour = "red", linetype = "dashed") +
    geom_text_repel(aes(label = sample)) +
    labs(x = "celltype proportion (%) at baseline", y = "virus RNA ratio (single dose/baseline)")
```

```{r}


select = c(0:8)[pvals$`Pr(>|t|)` < 0.08]
rsquare = data.frame(seurat_clusters = 5:6, c = c(cor(fdat$f[fdat$seurat_clusters == 5], fdat$ratio[fdat$seurat_clusters == 5]),
                                                  cor(fdat$f[fdat$seurat_clusters == 6], fdat$ratio[fdat$seurat_clusters == 6])))

ggplot(fdat[fdat$seurat_clusters %in% select,], aes(f, ratio)) + geom_point() + facet_wrap(~seurat_clusters) + 
  geom_abline(data = data.frame(seurat_clusters = 0:8, int = intercepts$Estimate, slope = pvals$Estimate)[select+1,],
              aes(intercept = int, slope = slope), colour = "darkblue", linetype = "dashed") +
    geom_label(data = data.frame(seurat_clusters = 5:6, l = paste("p =", round(pvals$`Pr(>|t|)`, 3)[select+1], "\nr2 =", round(rsquare$c,2))), 
              aes(label = l), x = 10, y = 1.1, colour = "darkblue") +
    theme_classic() +
    geom_hline(yintercept = 1) +
    geom_vline(data = divdata[divdata$seurat_clusters%in% select,], aes(xintercept = intercept), colour = "red", linetype = "dashed") +
    geom_text_repel(aes(label = donor_mapping[as.character(sample)])) +
    labs(x = "celltype proportion (%) at baseline", y = "virus RNA ratio (single dose/baseline)")->p;p

ggsave(p, filename = "../plots/Functional_composition_correlation2.pdf", width = 20, height = 9, units = "cm")
```

```{r}
m = FindAllMarkers(merged4, only.pos = F)

m %>% as_tibble() %>%
  filter(gene %in% row.names(topTags(lrt_rna_ratio, n = 50))) %>%
  mutate(direction = ifelse(avg_log2FC>0, "up","down")) %>%
  count(cluster, direction) %>%
  pivot_wider(names_from = direction, values_from = n) %>%
  mutate(ratio = log2(up/down)) %>%
  left_join(pvals %>% mutate(cluster = factor(0:11))) %>%
  ggplot(aes(ratio, Estimate)) + geom_point() +
  geom_text(aes(label = cluster), position = position_nudge(x=0.1, y=0.01)) +
  geom_hline(yintercept = 0, linetype="dashed")+
  geom_vline(xintercept = 0, linetype="dashed") +
  theme_classic()


```

```{r}
frequencies %>%
  filter(seurat_clusters %in% c(6,7,11,2,8,10), 
         !sample %in%  c("Donor 48", "Donor 54")) %>%
  mutate(dir = factor(ifelse(seurat_clusters %in% c(6,7,11), "negative", "positive"))) %>%
  ggplot(aes(reorder(seurat_clusters, as.numeric(dir)), f)) + 
  geom_bar(stat="identity", aes(fill = time), position = "dodge") +
  facet_wrap(~sample)

frequencies %>%
  filter(seurat_clusters %in% c(6,7,11,2,8,10), 
         !sample %in%  c("Donor 48", "Donor 54"), 
         time == "No dose") %>%
  inner_join(rna_ratios, by = c("sample" = "donor")) %>% 
  mutate(dir = factor(ifelse(seurat_clusters %in% c(6,7,11), "negative", "positive"))) %>%
  ggplot(aes(reorder(sample, ratio), f)) + 
  geom_bar(stat="identity") +
  facet_wrap(dir~seurat_clusters)
```

## Responder DEGs
```{r}
seu = merged4[,merged4$sample %in% names(donor_mapping)]
seu = SetIdent(seu, value = "seurat_clusters")
table(Idents(seu))
m5 = FindMarkers(seu, ident.1 = c("Donor 38", "Donor 49"), subset.ident = "5", group.by = "sample")
m6 = FindMarkers(seu, ident.1 = c("Donor 38", "Donor 49"), subset.ident = "6", group.by = "sample")
```

DEGs responders non responders
Figure 5
CD1-1
